### 2017-12-11 code.Rmd
first commit

% Code

## General approach

The data com from a series of different experiments. There is one single `xlsx` file with each experiment in a single sheet. A single `csv` file was created from exporting each of those sheets. We will conduct the analysis separated for each experiment. Follows the commented code used for these analyses separated per experiment.

## Load packages

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
library(readxl)
library(viridis)
library(ggthemes)
library(lme4)
library(cowplot)
library(broom)
library(agricolae)
library(lattice)
library(influence.ME)
```

## Perithecia production

### Import

Let's first import the data and have a look at the raw data.

```{r message=FALSE, warning=FALSE}
perithecia <- read_csv("./data/perithecia.csv") %>%
  filter(genotype != "t")

perithecia
```

### Transform
The data were prepared in the wide format where each score for percent perithecia coverage (0 -3) is in a separate colum. Let's transform this dataframe from the wide to the long format using `gather` function of `dplyr`. We need to create a column for the variable name and another for the score.

```{r}
perithecia2 <- perithecia %>%
  gather("score", "frequency", 8:11)

perithecia2
```

### Visualize

Let's have a look at the frequency of the scores for each species-genotype separated for each substrate. This is a good plot to show in the paper.

```{r}
g1 <- perithecia2 %>%
  ggplot(aes(reorder(genotype, -ppi), frequency, fill = score)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~substrate) +
  theme_few() + # load ggthemes
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "", y = "Proportion of grains", fill = "Score") +
  scale_fill_viridis(discrete = TRUE)
g1
```

### Model

Since we have one species which has two genotypes, it may be important, besides testing the effect of species or genotype in the model, to create a species*genotype interaction variable and fit a mixed model to test the effect of this variable on the perithecial production index (ppi), so that the two genotypes of one species can be compared.

```{r}
# creating the variable
perithecia <- perithecia %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F)
```

Now we fit the model using `lmer` function. We treat `Isolate` as a random effects.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
glmm_per <- lmer(
  ppi ~ species_genotype * substrate + (1 | Isolate),
  data = perithecia, REML = FALSE
)
```

Let's check the single and interaction effects.

```{r message=FALSE, warning=FALSE}
library(car)
Anova(glmm_per)
```

Evaluate the model

```{r}
library(lattice)
plot(glmm_per, type = c("p", "smooth"))
qqmath(glmm_per, id = 0.05)
```

#### Means comparison

Not too bad. Let's now compare the means of treatments and create a dataframe which will be use to further create a plot with the estimated means and confidence interval. We used the `emmeans` package which is an update for the old `lsmeans` package. The syntax is the same as before.

```{r message=FALSE, warning=FALSE}
library(emmeans)
medias <- emmeans(glmm_per, ~ species_genotype * substrate)
med <- cld(medias, Letters = LETTERS, alpha = .05)

# make output as a dataframe
med <- data.frame(med)
head(med)
```

### Figures

This new `med` dataframe does not contain the two original variables we want to use to plot the data. Let's separate the `species_genotypes` variable and create `genotype` and `species`.

```{r}
med2 <- med %>%
  separate(species_genotype, c("species", "genotype"), sep = "_", extra = "merge")
# extra argument was used to keep the genotype information alltogether
head(med2)
```

We now produce the plot with point estimate for the means and error bars for the 95%CI of the `ppi` for each `genotype`. We add `species` to the legend when using `color` argument.

```{r message=FALSE, warning=FALSE}
g2 <- med2 %>%
  ggplot(aes(reorder(genotype, -emmean), emmean, color = species)) +
  geom_point(position = position_dodge(width = 0.3)) +
  theme_few() +
  theme(legend.position = "right", legend.text = element_text(size = 6, face = "italic"), axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~substrate) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.2, position = position_dodge(width = 0.3)
  ) +
  labs(y = "Perithecia production index (%)", x = "Trichothecene genotype")
g2
```

Finally, we prepare and save the Figure 1 of the manuscript which is a combo figure with the the two previous plots for both the original scores and the normalized index. We use the `plot_grid` function of the `cowplot` package for this.

```{r message=FALSE, warning=FALSE}
grid1 <- plot_grid(g1, g2, labels = c("A", "B"), align = "hv", ncol = 1)
ggsave("figs/grid1.png", grid1, width = 6, height = 6)
```

## Mycelial growth

### Import

There is one `xlsx` file with data from a single experiment. We will use `read_excel` to correctly load the centigrades symbol for the temperature factors after setting locale to `pt_BR.UTF-8`.

```{r message=FALSE, warning=FALSE}
Sys.setlocale("LC_ALL", "pt_BR.UTF-8")
mgr <- read_excel("./data/mycelia.xlsx")

head(mgr)
```

### Visualize

We will now produce the plot with the means of mycelial growth rate for each species-genotype and temperature. This will be the final plot for publication.

```{r message=FALSE, warning=FALSE}
mgr1 <- mgr %>%
  group_by(isolate, genotype, species, temperature) %>%
  summarize(mean_mgr = mean(mgr)) %>%
  ggplot(aes(factor(genotype), mean_mgr, color = species)) +
  geom_jitter(width = 0.1, size = 3) +
  facet_wrap(~temperature) +
  scale_color_viridis(discrete = TRUE) +
  theme_few() +
  theme(
    legend.position = "none",
    legend.text = element_text(size = 7, face = "italic")
  ) +
  labs(
    x = "Trichothecene Genotype",
    y = "Mycelial growth rate (mm/day)", color = "Species"
  ) +
  ylim(0.2, 1.8)
```

### Model

Let's fit a mixed model for the `mgr` data. We first test the effect of the interaction.

```{r}
mgr <- mgr %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F)
lmer_mgr <- lmer(mgr ~ species_genotype * temperature + (1 | species_genotype / isolate), data = mgr, REML = FALSE)
Anova(lmer_mgr)
```

We now create different datasets for each temperature and test the effect of `species_genotype` within each temperature.

```{r}
# 15 C
mgr15 <- mgr %>%
  filter(temperature == "15 °C")

lmer_mgr15 <- lmer(mgr ~ species_genotype + (1 | species_genotype / isolate), data = mgr15, REML = FALSE)

Anova(lmer_mgr15)

# 30 C
mgr30 <- mgr %>%
  filter(temperature == "30 °C")

lmer_mgr30 <- lmer(mgr ~ species_genotype + (1 | species_genotype / isolate), data = mgr30, REML = FALSE)

Anova(lmer_mgr30)
```

We can see that there are no effect of `species_genotype` in the `mgr` within each of the temperatures.

## Sporoulation & Germination

In this experiment, the production of macroconidia and further germination of 20 randomly selected spores of the different strains of the species-genotype combination, were evaluated in two runs of the experiment.

### Import

```{r message=FALSE, warning=FALSE}
spor <- read_csv("./data/spor.csv")

# again, we create the species_genotype variable
spor <- spor %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F)
head(spor)
```

### Visualize

The code below will produce plots for total macroconidia production and percent germinated spores.

```{r message=FALSE, warning=FALSE}
library(scales)
spor1 <- spor %>%
  group_by(species, genotype, isolate) %>%
  summarize(mean_spor = mean(spores_ml))
spor2 <- spor1 %>%
  ggplot(aes(genotype, mean_spor, color = species)) +
  geom_jitter(width = 0.1, size = 3) +
  theme_few() +
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = c(1, 5), axis.text.x = element_text(angle = 0, hjust = 0.5, size = 7), legend.text = element_text(size = 7, face = "italic"), axis.title = element_text(size = 8)) +
  labs(x = "", y = "Macroconidia count (x 10.000)", color = "Species") +
  ylim(0, 5) +
  guides(colour = guide_legend(nrow = 2))
```

```{r message=FALSE, warning=FALSE}
germ1 <- spor %>%
  group_by(species, genotype, isolate) %>%
  summarize(mean_germ = mean(germp))
germ <- germ1 %>%
  ggplot(aes(genotype, mean_germ, color = species)) +
  geom_jitter(width = 0.1, size = 3) +
  theme_few() +
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust = 0.5, size = 7), legend.text = element_text(size = 7, face = "italic"), axis.title = element_text(size = 8)) +
  labs(x = "", y = "Germination rate (%)", color = "Species") +
  guides(colour = guide_legend(nrow = 2))
```

### Model

Anova model

```{r message=FALSE, warning=FALSE}
lmer_spor <- lmer(spores_ml ~ species_genotype + (1 | species_genotype), data = spor, REML = FALSE)
Anova(lmer_spor)
plot(lmer_spor, type = c("p", "smooth"))
qqmath(lmer_spor, id = 0.05)
```

#### Means comparison

```{r}
medias <- emmeans(lmer_spor, ~species_genotype)
med <- cld(medias, Letters = LETTERS, alpha = .05)
med <- data.frame(med)
med
```

### Figure

```{r message=FALSE, warning=FALSE}
grid1 <- plot_grid(spor2, germ, labels = c("B", "C"), ncol = 2, align = "hv")

grid5 <- plot_grid(mgr1, grid1, labels = c("A"), rel_heights = c(1, 1), ncol = 1, align = "hv")
ggsave("figs/mgr_spor_germ.png", grid5, width = 5, height = 5)

grid5
```

## Pathogenicity tests

### Import

```{r message=FALSE, warning=FALSE}
pathogen <- read_csv("./data/pathogenicity.csv")
pathogen <- pathogen %>%
  filter(isolate != "test")
```

### Transform

Let's prepare several summary statistics for the disease variables per cultivar. We will calculate the mean and standard deviation of proportion of diseased spikelets.

```{r}
# preparing cultivar 194

pathogen_194 <- pathogen %>%
  filter(cultivar == "BRS 194") %>%
  select(1:7, 9, 11, 12, 13) %>%
  gather(position, count, 8:9) %>%
  # filter(exp == 2) %>%
  group_by(cultivar, dai, genotype, species) %>%
  summarize(
    mean_sev = mean(sev),
    sd_sev = sd(sev)
  )

# preparing Guamirim

pathogen_Gua <- pathogen %>%
  filter(cultivar == "BRS Guamirim") %>%
  select(1:7, 9, 11:13) %>%
  gather(position, count, 8:9) %>%
  # filter(exp == 2) %>%
  group_by(cultivar, dai, genotype, species, position) %>%
  summarize(
    mean_sev = mean(sev),
    sd_sev = sd(sev)
  )
```

Let's calculate the area under the curve for each isolate. For this, we will use the `audpc` function of the `agricolae` package. First, we need to group several variables up to `dai` (days after inoculation). To make our life easier, lets use the `do` in conjuction with the `tidy` functions of the `dplyr` and `broom` package, respectively. This will calculate the audpce for each spike.

```{r}
audpc_194 <- pathogen %>%
  filter(cultivar == "BRS 194") %>%
  # filter(exp == 2) %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F) %>% 
  select(
    exp, cultivar, dai, isolate, species_genotype, species, genotype, spike,
    sev
  ) %>%
  group_by(exp, isolate, species_genotype, species, genotype, spike, dai) %>%
  summarize(mean_sev = mean(sev)) %>%
  do(tidy(audpc(.$mean_sev, .$dai)))
names(audpc_194)[8] <- "audpc"
```

```{r}
audpc_gua <- pathogen %>%
  filter(cultivar == "BRS Guamirim") %>%
  # filter(exp == 2) %>%
   unite(species_genotype, species, genotype, sep = "_", remove = F) %>% 
  select(
    exp, cultivar, dai, isolate, species_genotype, species, genotype, spike,
    sev
  ) %>%
  group_by(exp, isolate, species_genotype, species, genotype, spike, dai) %>%
  summarize(mean_sev = mean(sev)) %>%
  do(tidy(audpc(.$mean_sev, .$dai)))
names(audpc_gua)[8] <- "audpc"
```

### Visualize

```{r}

audpc_194 %>%
  group_by(species, genotype) %>%
  summarize(
    mean_audpc = mean(audpc),
    sd_audpc = sd(audpc)
  ) %>%
  ggplot(aes(species, mean_audpc, color = genotype)) +
  coord_flip() +
  theme_few() +
  geom_point(position = position_dodge(width = 0.4)) +
  geom_errorbar(aes(min = mean_audpc - sd_audpc, max = mean_audpc + sd_audpc), position = position_dodge(width = 0.4), width = 0.2)
```

```{r}
audpc_gua %>%
  group_by(species, genotype) %>%
  summarize(
    mean_audpc = mean(audpc),
    sd_audpc = sd(audpc)
  ) %>%
  ggplot(aes(species, mean_audpc, color = genotype)) +
  coord_flip() +
  theme_few() +
  geom_point(position = position_dodge(width = 0.4)) +
  geom_errorbar(aes(min = mean_audpc - sd_audpc, max = mean_audpc + sd_audpc), position = position_dodge(width = 0.4), width = 0.2)
```

### Model

#### BRS Guamirim

```{r}

mix_gua <- lmer(
  audpc ~ genotype +
    (1 | exp/spike),
  data = audpc_gua, REML = FALSE
)
```

Evaluate the model

```{r message=FALSE, warning=FALSE}

plot(mix_gua, type = c("p", "smooth"))
qqmath(mix_gua, id = 0.05)
```

```{r}
Anova(mix_gua)

library(emmeans)
means <- emmeans(mix_gua, ~ genotype)
cld(means)
```

```{r}
mix_194 <- lmer(log(audpc) ~ genotype + (1 | exp / spike), data = audpc_194, REML = FALSE)
```

Evaluate the model

```{r}

plot(mix_194, type = c("p", "smooth"))
qqmath(mix_194, id = 0.05)
```

#### BRS 194

```{r}
Anova(mix_194)
library(emmeans)
means <- emmeans(mix_194, ~ genotype)
cld(means)
```

### 2017-12-11 data.Rmd
first commit

% Data

## Raw data

The data were obtained from xx and were organized in xx excel/csv files. The example file is in the wide format where each day of assessment is in a different column, etc.

## Data wrangling

The original files were modified as following. The names of the variables were changed to.. The original wide format was retained and will be transformed using the dplyr functions, etc.

## Glimpse

```{r message=FALSE, warning=FALSE}
library(tidyverse)
#data <- read_delim("data/data.csv", delim = ";")
glimpse(data)
```

## Description of variables

- Variable1: 
  - Description 
  - Type: character
  - Unit

- Variable2: 
  - Description 
  - Type: integer
  - Unit

- Variable3: 
  - Description 
  - Type: double
  - Unit:

## Quick view

```{r}
data
```

## Download

The data are deposited in xx public repositories where they can be downloaded. Check on each data repository the policies regarding data use and citation.

<img src = "https://osf.io/static/img/circle_logo.png" width = "25">
[](DOI 10.17605/OSF.IO/7D3CH)

### 2017-12-11 index.Rmd
first commit

% About

<img src = "https://www.apsnet.org/edcenter/intropp/lessons/fungi/ascomycetes/PublishingImages/Fusariumdiscyclesm.jpg" align= right width = 250> This is a fully reproducible report that contains and describe the data and the R codes for analyzing and visualizing these data obtained from a series of experiments conducted as an MSc research in the Del Ponte lab. The research's objectives were to assess and compare members of a species complex (FGSC), with varying toxigenic potential, that cause [Fusarium head blight or scab](https://www.apsnet.org/edcenter/intropp/lessons/fungi/ascomycetes/Pages/Fusarium.aspx), an important disease of wheat and barley worldwide. Several fitness-related traits were compared in five experiments conducted both *in vitro* and in *planta*. Data from each experiment were analyzed separately.

## Summary of methods

A collection of 30 strains was obtained from the laboratory's FGSC collection and experiments were designed to measure traits related with asexual and sexual fungal growth, pathogenicity towards two wheat cultivars of varying susceptibility and sensitivity to tebuconazole fungicide. In most analyses, we were interested in differences among the species-genotype and not on individual isolates. Therefore, the isolates were treated as random sample representative of the population of strains. In the models, they were treated as random effects (mixed-modeling).

## Project team

Project PI:

- Emerson M. Del Ponte
- Universidade Federal de Viçosa

Members:

- Camila P. Nicolli, MSc student
- Piérri Spolti, Postdoctoral felow
- Franklin Machado, DSc student

## Citation

Please cite this research compendium as:

> Del Ponte EM. Title of compendium. Online at https://doi.org/xxx/xxx

## Licences

Data: [CC-0](https://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse
Manuscript: [CC-BY-4.0](https://creativecommons.org/licenses/by/4.0/)  
Code: [MIT](https://opensource.org/licenses/MIT) year: 2017, copyright holder: Emerson Del Ponte

## Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.html). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.html). By participating in this project you agree to abide by its terms.

### 2017-12-11 manuscript.Rmd
first commit

---
title: "My catchy title for a manuscript"
bibliography: crossref.bib
csl: plant-disease.csl
---

Author1, Author2, Author3

1 Affiliation 1
2 Affiliation 2

Corresponding author: Author 1

Article published in ...reference, link, DOI

# Abstract

Phasellus mattis mi et erat tristique, sit amet laoreet diam pulvinar. Sed rhoncus vulputate nibh, nec convallis metus viverra ultricies [@Lehner_2016]. Vivamus cursus feugiat ante, eget tincidunt dui tempus ac. Suspendisse a tempus leo. Curabitur varius neque finibus tincidunt aliquam. Orci varius natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Mauris eu ligula faucibus, ultricies risus vel, convallis mi [@Lehner_2016]. Aenean convallis vitae orci nec sollicitudin. Mauris mollis justo id vehicula elementum. Duis et neque eleifend, auctor ligula id, finibus ligula. Aliquam sagittis mauris a nisl consequat, eu mattis ante ultricies. Sed porttitor consequat placerat. Integer aliquam turpis nec tellus condimentum scelerisque. Nulla maximus metus ligula, sit amet sodales magna pellentesque nec. Cras accumsan pretium lectus a convallis.

Keywords: Aliquam sagittis mauris

# Introduction

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Ut vehicula risus nec libero imperdiet fermentum. Maecenas convallis libero condimentum finibus euismod. Nunc lacinia, leo et sagittis posuere, augue sapien molestie lacus, ut imperdiet velit sem et nunc. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam euismod lorem vitae consequat congue. Donec bibendum, nulla et venenatis tincidunt, nulla quam hendrerit lorem, nec semper nunc tortor sit amet dui. Maecenas sit amet laoreet sem.

Phasellus mattis mi et erat tristique, sit amet laoreet diam pulvinar. Sed rhoncus vulputate nibh, nec convallis metus viverra ultricies. Vivamus cursus feugiat ante, eget tincidunt dui tempus ac. Suspendisse a tempus leo. Curabitur varius neque finibus tincidunt aliquam. Orci varius natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Mauris eu ligula faucibus, ultricies risus vel, convallis mi. Aenean convallis vitae orci nec sollicitudin. Mauris mollis justo id vehicula elementum. Duis et neque eleifend, auctor ligula id, finibus ligula. Aliquam sagittis mauris a nisl consequat, eu mattis ante ultricies. Sed porttitor consequat placerat. Integer aliquam turpis nec tellus condimentum scelerisque. Nulla maximus metus ligula, sit amet sodales magna pellentesque nec. Cras accumsan pretium lectus a convallis.

Aliquam erat volutpat. Cras commodo eros arcu, ac pellentesque augue gravida at. Nulla feugiat dui nec ligula aliquam, ut lacinia ex facilisis. Nunc ut metus pulvinar, pulvinar sem vel, maximus lectus. Nulla commodo lorem vitae porttitor tempus. Nullam sed nibh sed eros aliquet congue. Proin sollicitudin vestibulum ex et tincidunt.

Nulla rutrum leo nec feugiat fringilla. Suspendisse sed hendrerit orci. Fusce quis lobortis nibh, eget luctus mi. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Nulla nec purus a lectus semper congue sit amet quis ipsum. Mauris accumsan dapibus auctor. Curabitur rhoncus massa eu felis ultricies mattis vel vel sem.

Aliquam sit amet lacinia massa, nec convallis ex. Nullam porttitor tincidunt mauris, elementum pellentesque velit tristique nec. Nunc a sollicitudin ipsum. Phasellus felis turpis, pretium vel scelerisque eget, ultricies id ex. Mauris in ullamcorper felis. Quisque nec malesuada urna. Integer pellentesque, magna vitae imperdiet ullamcorper, augue lacus blandit tortor, quis vulputate dolor elit non elit. Nullam euismod orci et rhoncus fringilla. Quisque vitae mauris sed eros sagittis lacinia. Vestibulum ipsum lorem, pretium a orci et, sagittis ultrices mi. Ut id eleifend libero. Fusce est ipsum, finibus id quam id, auctor tincidunt tortor. Suspendisse accumsan sit amet sapien at rutrum.

# Material and Methods

## item1

Cras sollicitudin massa et felis rhoncus, scelerisque feugiat metus rutrum. Aenean blandit augue id sapien mattis consectetur. Nulla sollicitudin lectus nec massa iaculis, sed aliquet ante consequat. Donec ut vehicula felis. Etiam placerat convallis efficitur. Mauris id lacinia est. Vivamus vitae diam metus. Aliquam eget massa at ante ornare ultrices id nec tortor. Curabitur quam odio, tincidunt sit amet ipsum id, hendrerit viverra erat. Integer lacinia eros nec metus ultrices, vitae semper sapien rhoncus. Nulla id tempor ante. Nullam sem turpis, sagittis nec risus sed, imperdiet scelerisque lacus. Vivamus tincidunt felis a risus iaculis lacinia. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.

## item2
Vivamus ullamcorper justo sapien, vitae efficitur dui laoreet vitae. Etiam tristique, nibh vitae varius mattis, turpis leo elementum eros, quis dictum dolor enim a nunc. Etiam cursus velit eget nisl gravida, quis faucibus justo congue. Nunc vestibulum venenatis felis. Nullam sit amet nibh erat. Nulla vel ex viverra, fringilla ex eget, luctus enim. Nulla lacinia nisi risus, sit amet laoreet purus mattis quis. Aliquam pellentesque dolor ac metus blandit iaculis. Nulla mollis massa magna, non auctor nibh tempor vitae. Quisque mollis vel dolor eu pretium.

## item3

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Proin in blandit est. Fusce quis hendrerit libero, id auctor quam. Duis et vestibulum mauris. Maecenas sagittis condimentum laoreet. Donec vitae lectus eu erat hendrerit facilisis sed eget sem. Nam interdum libero ac ex cursus suscipit.

# Results

## item1

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Proin in blandit est. Fusce quis hendrerit libero, id auctor quam. Duis et vestibulum mauris. Maecenas sagittis condimentum laoreet. Donec vitae lectus eu erat hendrerit facilisis sed eget sem. Nam interdum libero ac ex cursus suscipit.

## item2

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Proin in blandit est. Fusce quis hendrerit libero, id auctor quam. Duis et vestibulum mauris. Maecenas sagittis condimentum laoreet. Donec vitae lectus eu erat hendrerit facilisis sed eget sem. Nam interdum libero ac ex cursus suscipit.

## item3

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Proin in blandit est. Fusce quis hendrerit libero, id auctor quam. Duis et vestibulum mauris. Maecenas sagittis condimentum laoreet. Donec vitae lectus eu erat hendrerit facilisis sed eget sem. Nam interdum libero ac ex cursus suscipit.

# Discussion

Sed vitae neque diam. Donec facilisis nisi nec mollis finibus. Nullam fermentum, nisl luctus vehicula vestibulum, risus eros molestie enim, eget auctor nunc libero sit amet nulla. Suspendisse ac consectetur sapien. Sed lacinia, neque lacinia blandit congue, sapien erat malesuada augue, ac vestibulum quam metus id odio. Morbi metus magna, tempor quis enim id, commodo eleifend risus. Sed quis mollis lectus, sed imperdiet tortor. Pellentesque eget condimentum velit, a imperdiet ante. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae;

Vivamus ullamcorper justo sapien, vitae efficitur dui laoreet vitae. Etiam tristique, nibh vitae varius mattis, turpis leo elementum eros, quis dictum dolor enim a nunc. Etiam cursus velit eget nisl gravida, quis faucibus justo congue. Nunc vestibulum venenatis felis. Nullam sit amet nibh erat. Nulla vel ex viverra, fringilla ex eget, luctus enim. Nulla lacinia nisi risus, sit amet laoreet purus mattis quis. Aliquam pellentesque dolor ac metus blandit iaculis. Nulla mollis massa magna, non auctor nibh tempor vitae. Quisque mollis vel dolor eu pretium.

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Proin in blandit est. Fusce quis hendrerit libero, id auctor quam. Duis et vestibulum mauris. Maecenas sagittis condimentum laoreet. Donec vitae lectus eu erat hendrerit facilisis sed eget sem. Nam interdum libero ac ex cursus suscipit.

Integer efficitur viverra elementum. Quisque commodo blandit blandit. Donec eget ex et ante iaculis porta. In sem augue, lacinia ut venenatis ullamcorper, volutpat quis tortor. Proin molestie ultrices purus ac consectetur. Duis ultrices, nunc vel euismod finibus, dui nunc elementum mauris, cursus vestibulum lorem magna in dui. Vivamus euismod consequat dolor.

Cras id sem augue. Nullam tincidunt vulputate arcu in dignissim. Vivamus gravida feugiat turpis at cursus. Donec euismod sit amet orci vel faucibus. In convallis rutrum leo eget elementum. Fusce mattis lacus quis risus volutpat, vel imperdiet ex dapibus. Quisque imperdiet est ac eros dapibus, dapibus aliquam justo porttitor.

# Acknowledgements

 Fusce quis hendrerit libero, id auctor quam. Duis et vestibulum mauris. Maecenas sagittis condimentum laoreet. Donec vitae lectus eu erat hendrerit facilisis sed eget sem. Nam interdum libero ac ex cursus suscipit.

# References

### 2017-12-12 code.Rmd
updates on paths

@@
 library(tidyverse)
+library(here)
 library(readxl)
@@
 ```{r message=FALSE, warning=FALSE}
-perithecia <- read_csv("./data/perithecia.csv") %>%
+perithecia <- read_excel(here("data", "all-data.xlsx"), "perithecia") %>%
   filter(genotype != "t")
@@
 ### Transform
+
 The data were prepared in the wide format where each score for percent perithecia coverage (0 -3) is in a separate colum.
@@
 grid1 <- plot_grid(g1, g2, labels = c("A", "B"), align = "hv", ncol = 1)
-ggsave("figs/grid1.png", grid1, width = 6, height = 6)
+ggsave(here("figs", "figure1.png"), grid1, width = 6, height = 6)
+grid1
 ```
@@
 Sys.setlocale("LC_ALL", "pt_BR.UTF-8")
-mgr <- read_excel("./data/mycelia.xlsx")
+mgr <- read_excel(here("data", "all-data.xlsx"), "mycelia")
 
@@
 ```{r message=FALSE, warning=FALSE}
-spor <- read_csv("./data/spor.csv")
+spor <- read_excel(here("data", "all-data.xlsx"), "spor_germ")
 
@@
 
-### Figure
+### Figure 3
 
@@
 grid5 <- plot_grid(mgr1, grid1, labels = c("A"), rel_heights = c(1, 1), ncol = 1, align = "hv")
-ggsave("figs/mgr_spor_germ.png", grid5, width = 5, height = 5)
+ggsave(here("figs", "figure3.png"), grid5, width = 5, height = 5)
 
@@
 ```{r message=FALSE, warning=FALSE}
-pathogen <- read_csv("./data/pathogenicity.csv")
-pathogen <- pathogen %>%
+pathogen <- read_excel(here("data", "all-data.xlsx"), "pathogenicity") %>%
   filter(isolate != "test")
@@
 ```
+
+## Fungicide sensitivity
+
+### Import
+
+```{r message=FALSE, warning=FALSE}
+ec50 <- read_excel(here("data", "all-data.xlsx"), "ec50") 
+```
+
+### Summarize
+
+```{r}
+ec502 <- ec50 %>%
+  group_by(isolate, species, genotype)%>%
+  summarize(mean_ec50 = mean(ec50),
+            sd_ec50 = sd(ec50)
+            )
+```
+
+### Visualize
+
+Let's have a look at the mean EC50 for the isolates
+
+```{r}
+
+library(viridis)
+library(ggthemes)
+ec502 %>%
+  ggplot(aes(genotype, mean_ec50, color = species))+
+  geom_jitter(width = 0.1)+
+  #geom_errorbar(aes(min = mean_ec50 - sd_ec50, max = mean_ec50 + sd_ec50), width=0.2, position = position_dodge(width = 0.2))+
+  theme_few()+
+  theme(legend.position="right", legend.text=element_text(size=7, face = 'italic'))+
+    ylim(0,1.8)+
+ scale_color_viridis( discrete = TRUE)+
+  labs(y = (expression(paste('EC'[50], ' (', mu,'g/ml)'))), x = "Trichothecene genotype", color="Species")+
+  ggsave(here("figs", "ec50-new.png"), width=4.5, height=3)
+
+```

### 2017-12-12 data.Rmd
updates on paths

@@
 
-The data were obtained from xx and were organized in xx excel/csv files. The example file is in the wide format where each day of assessment is in a different column, etc.
+The data were obtained from five different, mostly repeated experiments, and were organized in a single `xlsx` file.
+Data from each experiment are in a separate sheet, which are loaded separately using the `read_excel` function of the `readxl` file.
 
-## Data wrangling
+`I found much more convenient to use this binary file instead of storing in a text file like csv.
+For some variabeles, I had to deal with some characters which were not loading correctly when using `csv` files.`
 
-The original files were modified as following. The names of the variables were changed to.. The original wide format was retained and will be transformed using the dplyr functions, etc.
-
-## Glimpse
-
-```{r message=FALSE, warning=FALSE}
-library(tidyverse)
-#data <- read_delim("data/data.csv", delim = ";")
-glimpse(data)
-```
-
-## Description of variables
-
-- Variable1: 
-  - Description 
-  - Type: character
-  - Unit
-
-- Variable2: 
-  - Description 
-  - Type: integer
-  - Unit
-
-- Variable3: 
-  - Description 
-  - Type: double
-  - Unit:
-
-## Quick view
-
-```{r}
-data
-```
+Most of the data are already organized in the long format, but some are in the wide, and so will need some transformation. These are done within the respective section of the data analysis page.
 
@@
 
-The data are deposited in xx public repositories where they can be downloaded. Check on each data repository the policies regarding data use and citation.
+The data are stored in the `data` folder of the GitHub repository.
 
-<img src = "https://osf.io/static/img/circle_logo.png" width = "25">
-[](DOI 10.17605/OSF.IO/7D3CH)
+https://github.com/phytopathology/paper-FGSC-fitness

### 2017-12-12 code.Rmd
text changes

@@
 
-## General approach
+## Introduction
 
@@
 There is one single `xlsx` file with each experiment in a single sheet.
-A single `csv` file was created from exporting each of those sheets.
-We will conduct the analysis separated for each experiment.
 Follows the commented code used for these analyses separated per experiment.
@@
 
-The data were prepared in the wide format where each score for percent perithecia coverage (0 -3) is in a separate colum.
+The data were prepared in the wide format where each score for percent perithecia coverage (0 -3) is in a separate column.
 Let's transform this dataframe from the wide to the long format using `gather` function of `dplyr`.
@@
 
-Let's have a look at the frequency of the scores for each species-genotype separated for each substrate. This is a good plot to show in the paper.
+Let's visualize the frequency of the scores for each species-genotype separated for each substrate. This is a nice plot to show in the paper!
 
@@
 
-Since we have one species which has two genotypes, it may be important, besides testing the effect of species or genotype in the model, to create a species*genotype interaction variable and fit a mixed model to test the effect of this variable on the perithecial production index (ppi), so that the two genotypes of one species can be compared.
+Since we have one species which has two genotypes, it may be important, besides testing the effect of species or genotype in the model, to create a species*genotype interaction variable and fit a mixed model to test the effect of this variable on the perithecial production index (ppi).
+This way the two genotypes within one species can be compared between them and with the others.
 

### 2017-12-12 data.Rmd
text changes

@@
 
-The data were obtained from five different, mostly repeated experiments, and were organized in a single `xlsx` file.
-Data from each experiment are in a separate sheet, which are loaded separately using the `read_excel` function of the `readxl` file.
+The data were obtained from five different (mostly repeated in time) experiments, and were organized in a single `xlsx` file.
+Data from each experiment are in a separate sheet, which are loaded separately by specifying the respective sheet name using the `read_excel` function of the `readxl` file.
 
-`I found much more convenient to use this binary file instead of storing in a text file like csv.
-For some variabeles, I had to deal with some characters which were not loading correctly when using `csv` files.`
+**NOTE:** Although text files are prefered to store the data, I found more convenient to use a binary file instead of storing in a text file like csv.
+For some variabeles, I had to deal with some characters which were not loading correctly when using `csv` files.
+To visualize the data, go to the respective section of the data analysis page.
 
-Most of the data are already organized in the long format, but some are in the wide, and so will need some transformation. These are done within the respective section of the data analysis page.
+Most of the data are already organized in a rectangular (tidy) format, mostly in the long-format (a raw is an observation), but some are in the wide (observations in multiple colums), and so will need some wrangling.
+This munging is done within the respective section of the data analysis page, after data import.
 

### 2017-12-12 index.Rmd
text changes

@@
 
-<img src = "https://www.apsnet.org/edcenter/intropp/lessons/fungi/ascomycetes/PublishingImages/Fusariumdiscyclesm.jpg" align= right width = 250> This is a fully reproducible report that contains and describe the data and the R codes for analyzing and visualizing these data obtained from a series of experiments conducted as an MSc research in the Del Ponte lab.
-The research's objectives were to assess and compare members of a species complex (FGSC), with varying toxigenic potential, that cause [Fusarium head blight or scab](https://www.apsnet.org/edcenter/intropp/lessons/fungi/ascomycetes/Pages/Fusarium.aspx), an important disease of wheat and barley worldwide.
+<img src = "https://www.apsnet.org/edcenter/intropp/lessons/fungi/ascomycetes/PublishingImages/Fusariumdiscyclesm.jpg" align= right width = 250> This is a research compendium that contains and describe the data and the R codes for analyzing and visualizing these data obtained from a series of experiments conducted as an MSc research in the Del Ponte lab.
+The main research's objective was to assess and compare members of a species complex (FGSC), with varying toxigenic potential, that cause [Fusarium head blight or scab](https://www.apsnet.org/edcenter/intropp/lessons/fungi/ascomycetes/Pages/Fusarium.aspx), an important disease of wheat and barley worldwide.
 Several fitness-related traits were compared in five experiments conducted both *in vitro* and in *planta*.
-Data from each experiment were analyzed separately.
+Check the manuscript page.
 
@@
 A collection of 30 strains was obtained from the laboratory's FGSC collection and experiments were designed to measure traits related with asexual and sexual fungal growth, pathogenicity towards two wheat cultivars of varying susceptibility and sensitivity to tebuconazole fungicide.
-In most analyses, we were interested in differences among the species-genotype and not on individual isolates.
+In most analyses, we were interested in testing the effect of species or toxing genotypes, and not the effect of individual isolates.
 Therefore, the isolates were treated as random sample representative of the population of strains.
-In the models, they were treated as random effects (mixed-modeling).
+We used a mixed modeling approach where isolates were treated as random effects.
+Data from each experiment were analyzed separately.
 
@@
 
-> Del Ponte EM. Title of compendium. Online at https://doi.org/xxx/xxx
+> Del Ponte EM. Fitness-related traits of Fusarium graminearum species complex: a research compendia. Online at https://phytopathology.github.io/paper-FGSC-fitness/
 
@@
 Code: [MIT](https://opensource.org/licenses/MIT) year: 2017, copyright holder: Emerson Del Ponte
-
-## Contributions
-
-We welcome contributions from everyone.
-Before you get started, please see our [contributor guidelines](CONTRIBUTING.html).
-Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.html).
-By participating in this project you agree to abide by its terms.

### 2017-12-12 code.Rmd
new figures

@@
 library(lattice)
-library(influence.ME)
+library(emmeans)
+library(car)
+library(scales)
 ```
@@
 ```{r message=FALSE, warning=FALSE}
-library(car)
 Anova(glmm_per)
@@
 ```{r}
-library(lattice)
 plot(glmm_per, type = c("p", "smooth"))
@@
 ```{r message=FALSE, warning=FALSE}
-library(emmeans)
 medias <- emmeans(glmm_per, ~ species_genotype * substrate)
@@
 
-This new `med` dataframe does not contain the two original variables we want to use to plot the data. Let's separate the `species_genotypes` variable and create `genotype` and `species`.
+This new `med` dataframe does not contain the two original variables we want to use to plot the data. Let's separate the `species_genotype` variable and create `genotype` and `species`.
 
@@
 ```{r message=FALSE, warning=FALSE}
-Sys.setlocale("LC_ALL", "pt_BR.UTF-8")
+
 mgr <- read_excel(here("data", "all-data.xlsx"), "mycelia")
@@
 mgr15 <- mgr %>%
-  filter(temperature == "15 °C")
+  filter(temperature == "15")
 
@@
 mgr30 <- mgr %>%
-  filter(temperature == "30 °C")
+  filter(temperature == "30")
 
@@
 ```{r message=FALSE, warning=FALSE}
-library(scales)
 spor1 <- spor %>%
@@
 grid5 <- plot_grid(mgr1, grid1, labels = c("A"), rel_heights = c(1, 1), ncol = 1, align = "hv")
-ggsave(here("figs", "figure3.png"), grid5, width = 5, height = 5)
+ggsave(here("figs", "figure2.png"), grid5, width = 5, height = 5)
 
@@
 
+### Visualize temporal
+
+```{r message=FALSE, warning=FALSE}
+
+brs194 <- pathogen_194 %>%
+  group_by(dai, genotype, species) %>%
+  ggplot(aes(dai, mean_sev, color = species, shape = genotype, group = interaction(species, genotype))) +
+  geom_point(position = position_dodge(width = 0.9)) +
+  geom_line() +
+  geom_errorbar(
+    aes(min = mean_sev - sd_sev, max = mean_sev + sd_sev),
+    width = 0.2, alpha = 0.3, position = position_dodge(width = 0.9)
+  ) +
+  theme_few() +
+  ylim(0, 55) +
+  xlim(0, 25) +
+  theme(legend.position = "right", legend.text = element_text(size = 7, face = "italic")) +
+  scale_color_viridis(discrete = TRUE) +
+  labs(
+    y = "FHB severity (%) ", x = "Days after inoculation",
+    color = "species"
+  )
+
+brsgua <- pathogen_Gua %>%
+  group_by(dai, genotype, species) %>%
+  ggplot(aes(dai, mean_sev, color = species, shape = genotype, group = interaction(species, genotype))) +
+  geom_point(position = position_dodge(width = 0.9)) +
+  geom_line() +
+  geom_errorbar(
+    aes(min = mean_sev - sd_sev, max = mean_sev + sd_sev),
+    width = 0.2, alpha = 0.3, position = position_dodge(width = 0.9)
+  ) +
+  theme_few() +
+  theme(legend.position = "none") +
+  scale_color_viridis(discrete = TRUE) +
+  ylim(0, 55) +
+  xlim(0, 25) +
+  labs(
+    y = "FHB severity (%) ", x = "Days after inoculation",
+    color = "species"
+  )
+
+fig3 <- plot_grid(brsgua, brs194, labels = c("A", "B"), rel_widths = c(0.7, 1), ncol = 2, align = "hv")
+ggsave(here("figs", "figure3.png"), fig3, width = 9, height = 4)
+
+fig3
+```
+
+### Visualize AUDPC
+
 Let's calculate the area under the curve for each isolate.
@@
 
-### Visualize
-
 ```{r}
-
 audpc_194 %>%
@@
   audpc ~ genotype +
-    (1 | exp/spike),
+    (1 | exp / spike),
   data = audpc_gua, REML = FALSE
@@
 
-library(emmeans)
 means <- emmeans(mix_gua, ~ genotype)
@@
 Anova(mix_194)
-library(emmeans)
 means <- emmeans(mix_194, ~ genotype)
@@
 ec502 <- ec50 %>%
-  group_by(isolate, species, genotype)%>%
-  summarize(mean_ec50 = mean(ec50),
+  group_by(isolate, species, genotype) %>%
+  summarize(
+    mean_ec50 = mean(ec50),
     sd_ec50 = sd(ec50)
@@
 ```{r}
-
-library(viridis)
-library(ggthemes)
 ec502 %>%
-  ggplot(aes(genotype, mean_ec50, color = species))+
-  geom_jitter(width = 0.1)+
-  #geom_errorbar(aes(min = mean_ec50 - sd_ec50, max = mean_ec50 + sd_ec50), width=0.2, position = position_dodge(width = 0.2))+
-  theme_few()+
-  theme(legend.position="right", legend.text=element_text(size=7, face = 'italic'))+
-    ylim(0,1.8)+
- scale_color_viridis( discrete = TRUE)+
-  labs(y = (expression(paste('EC'[50], ' (', mu,'g/ml)'))), x = "Trichothecene genotype", color="Species")+
-  ggsave(here("figs", "ec50-new.png"), width=4.5, height=3)
-
+  ggplot(aes(genotype, mean_ec50, color = species)) +
+  geom_jitter(width = 0.1) +
+  # geom_errorbar(aes(min = mean_ec50 - sd_ec50, max = mean_ec50 + sd_ec50), width=0.2, position = position_dodge(width = 0.2))+
+  theme_few() +
+  theme(legend.position = "right", legend.text = element_text(size = 7, face = "italic")) +
+  ylim(0, 1.8) +
+  scale_color_viridis(discrete = TRUE) +
+  labs(y = (expression(paste("EC"[50], " (", mu, "g/ml)"))), x = "Trichothecene genotype", color = "Species") +
+  ggsave(here("figs", "ec50-new.png"), width = 4.5, height = 3)
 ```

### 2017-12-16 data.Rmd
Updated manuscript

--- 2017-12-12 data.Rmd
@@
 
-The data were obtained from five different, mostly repeated experiments, and were organized in a single `xlsx` file.
-Data from each experiment are in a separate sheet, which are loaded separately using the `read_excel` function of the `readxl` file.
+The data were obtained from five different (mostly repeated in time) experiments, and were organized in a single `xlsx` file.
+Data from each experiment are in a separate sheet and were exported to single `csv` files.
 
-`I found much more convenient to use this binary file instead of storing in a text file like csv.
-For some variabeles, I had to deal with some characters which were not loading correctly when using `csv` files.`
-
-Most of the data are already organized in the long format, but some are in the wide, and so will need some transformation. These are done within the respective section of the data analysis page.
+Most of the data are already organized in a rectangular (tidy) shape and long-format (a raw is an observation).
+Howver, some are in the wide (observations in multiple colums), and these will require some wrangling.
+Data munging is done within the respective section of the data analysis page, after data import.
 

### 2017-12-16 code.Rmd
new changes

@@
 library(here)
-library(readxl)
 library(viridis)
@@
 ```{r message=FALSE, warning=FALSE}
-perithecia <- read_excel(here("data", "all-data.xlsx"), "perithecia") %>%
+perithecia <- read_csv(here("data", "perithecia.csv")) %>%
   filter(genotype != "t")
@@
   theme_few() +
-  theme(legend.position = "right", legend.text = element_text(size = 6, face = "italic"), axis.text.x = element_text(angle = 45, hjust = 1)) +
+  theme(legend.position = "right", legend.text = element_text(size = 8, face = "italic"), axis.text.x = element_text(angle = 45, hjust = 1)) +
   scale_color_viridis(discrete = TRUE) +
@@
 grid1 <- plot_grid(g1, g2, labels = c("A", "B"), align = "hv", ncol = 1)
-ggsave(here("figs", "figure1.png"), grid1, width = 6, height = 6)
+ggsave(here("figs", "figure1.png"), grid1, width = 6, height = 7)
 grid1
@@
 
-There is one `xlsx` file with data from a single experiment. We will use `read_excel` to correctly load the centigrades symbol for the temperature factors after setting locale to `pt_BR.UTF-8`.
+There is one `csv` file with data from a single experiment. We will use `read_csv` to impor the data.
 
@@
 
-mgr <- read_excel(here("data", "all-data.xlsx"), "mycelia")
+mgr <- read_csv(here("data", "mycelia.csv"))
 
@@
     x = "Trichothecene Genotype",
-    y = "Mycelial growth rate (mm/day)", color = "Species"
+    y = "MGR (mm/day)", color = "Species"
   ) +
@@
 ```{r message=FALSE, warning=FALSE}
-spor <- read_excel(here("data", "all-data.xlsx"), "spor_germ")
+spor <- read_csv(here("data", "spor_germ.csv"))
 
@@
   scale_color_viridis(discrete = TRUE) +
-  theme(legend.position = c(1, 5), axis.text.x = element_text(angle = 0, hjust = 0.5, size = 7), legend.text = element_text(size = 7, face = "italic"), axis.title = element_text(size = 8)) +
-  labs(x = "", y = "Macroconidia count (x 10.000)", color = "Species") +
+  theme(legend.position = c(1, 5), axis.text.x = element_text(angle = 0, hjust = 0.5, size = 7), legend.text = element_text(size = 12, face = "italic"), axis.title = element_text(size = 12)) +
+  labs(x = "", y = "N. of macroconidia (x10.000)", color = "Species") +
   ylim(0, 5) +
@@
   scale_color_viridis(discrete = TRUE) +
-  theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust = 0.5, size = 7), legend.text = element_text(size = 7, face = "italic"), axis.title = element_text(size = 8)) +
+  theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust = 0.5, size = 7), legend.text = element_text(size = 12, face = "italic"), axis.title = element_text(size = 12)) +
   labs(x = "", y = "Germination rate (%)", color = "Species") +
@@
 
-### Figure 3
+### Figure 2
 
@@
 grid5 <- plot_grid(mgr1, grid1, labels = c("A"), rel_heights = c(1, 1), ncol = 1, align = "hv")
-ggsave(here("figs", "figure2.png"), grid5, width = 5, height = 5)
+ggsave(here("figs", "figure2.png"), grid5, width = 6, height = 6.2)
 
@@
 ```{r message=FALSE, warning=FALSE}
-pathogen <- read_excel(here("data", "all-data.xlsx"), "pathogenicity") %>%
+pathogen <- read_csv(here("data", "pathogenicity.csv")) %>%
   filter(isolate != "test")
@@
   xlim(0, 25) +
-  theme(legend.position = "right", legend.text = element_text(size = 7, face = "italic")) +
+  theme(legend.position = "right", legend.text = element_text(size = 9, face = "italic")) +
   scale_color_viridis(discrete = TRUE) +
@@
 
-plot(mix_gua, type = c("p", "smooth"))
-qqmath(mix_gua, id = 0.05)
+plot(mix_gua, type = c("p", "smooth"), col = "black")
+qqmath(mix_gua, id = 0.05, col = "black")
 ```
@@
 
-plot(mix_194, type = c("p", "smooth"))
-qqmath(mix_194, id = 0.05)
+plot(mix_194, type = c("p", "smooth"), col = "black")
+qqmath(mix_194, id = 0.05, col = "black")
 ```
@@
 ```{r message=FALSE, warning=FALSE}
-ec50 <- read_excel(here("data", "all-data.xlsx"), "ec50")
+ec50 <- read_csv(here("data", "ec50.csv"))
 ```

### 2017-12-16 data.Rmd
new changes

@@
 The data were obtained from five different (mostly repeated in time) experiments, and were organized in a single `xlsx` file.
-Data from each experiment are in a separate sheet and were exported to single `csv` files.
+Data from each experiment are in a separate sheet and were exported to a `csv` file.
 
-Most of the data are already organized in a rectangular (tidy) shape and long-format (a raw is an observation).
-Howver, some are in the wide (observations in multiple colums), and these will require some wrangling.
-Data munging is done within the respective section of the data analysis page, after data import.
+The data are already organized in a rectangular (tidy) format, mostly in the long-format (a raw is an observation), but some are in the wide (observations in multiple colums), and so will need some wrangling.
+This munging is done within the respective section of the data analysis page, after data import.
 
@@
 
-https://github.com/phytopathology/paper-FGSC-fitness
+https://github.com/phytopathology/paper-FGSC-fitness/data

### 2017-12-16 manuscript.Rmd

@@
 ---
-title: "My catchy title for a manuscript"
+title: "Fitness traits of deoxynivalenol and nivalenol-producing Fusarium graminearum species complex strains from wheat"
 bibliography: crossref.bib
@@
 
-Author1, Author2, Author3
+Camila P. Nicolli, Franklin J. Machado, Piérri Spolti, Emerson M. Del Ponte   
+Departamento de Fitopatologia, Universidade Federal de Viçosa, 36570-000, Viçosa, MG, Brazil.
 
-1 Affiliation 1
-2 Affiliation 2
+Corresponding author: Emerson M. Del Ponte, email: delponte@ufv.br
 
-Corresponding author: Author 1
+## Abstract
 
-Article published in ...reference, link, DOI
+> *Fusarium graminearum* of the 15-acetyl(A)deoxynivalenol(DON) chemotype is the main cause of Fusarium head blight (FHB) of wheat in south of Brazil.
+However, 3-ADON and nivalenol(NIV) chemotypes have been found in other members of the species complex causing FHB in wheat.
+To improve our understanding of the pathogen ecology, we assessed a range of fitness-related traits in a sample of 30 strains representatives of 15-ADON (**F.
+graminearum**), 3-ADON (*F.
+cortaderiae* and *F.
+austroamericanum*) and NIV (*F.
+meridionale* and *F.
+cortaderiae*).
+These included: perithecia formation on three cereal-based substrates, mycelial growth at two suboptimal temperatures, sporulation and germination, pathogenicity towards a susceptible and a moderately resistant cultivar and sensitivity to tebuconazole.
+The most important trait favoring *F.
+graminearum* was its 2x higher sexual fertility (> 40% PPI = perithecia production index) than the other species (< 30% PPI); PPI varied among substrates (maize > rice > wheat).
+In addition, sensitivity to tebuconazole appeared lower in *F.
+graminearum* which had the only strain with EC50 > 1 ppm.
+In the pathogenicity assays, the DON-producers were generally more aggressive (1.5 to 2x higher final severity) towards the two cultivars, with 3-ADON or 15-ADON leading to higher area under the severity curve than the NIV strains in the susceptible and moderately resistant cv., respectively.
+There was significant variation among strains of a same species with regards asexual fertility (mycelial growth, macroconidia production and germination), which suggest a strain- rather than a species-specific differences.
+These results contribute new knowledge to improve our understanding of the pathogen-related traits that may explain the dominance of certain members of the species complex in specific wheat agroecosystems.
 
-# Abstract
+Key words: Triticum aestivum, trichothecenes, fitness, FGSC.
 
-Phasellus mattis mi et erat tristique, sit amet laoreet diam pulvinar.
-Sed rhoncus vulputate nibh, nec convallis metus viverra ultricies [@Lehner_2016].
-Vivamus cursus feugiat ante, eget tincidunt dui tempus ac.
-Suspendisse a tempus leo.
-Curabitur varius neque finibus tincidunt aliquam.
-Orci varius natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
-Mauris eu ligula faucibus, ultricies risus vel, convallis mi [@Lehner_2016].
-Aenean convallis vitae orci nec sollicitudin.
-Mauris mollis justo id vehicula elementum.
-Duis et neque eleifend, auctor ligula id, finibus ligula.
-Aliquam sagittis mauris a nisl consequat, eu mattis ante ultricies.
-Sed porttitor consequat placerat.
-Integer aliquam turpis nec tellus condimentum scelerisque.
-Nulla maximus metus ligula, sit amet sodales magna pellentesque nec.
-Cras accumsan pretium lectus a convallis.
+## Introduction
 
-Keywords: Aliquam sagittis mauris
+Fusarium head blight (FHB) is a major disease of wheat and small grain crops due to the reduction in yield and contamination of grain with dangerous mycotoxins (McMullen et al.
+2012).
+Deoxynivalenol (DON) is the most significant mycotoxin given its widespread occurrence at levels of concern to human and animal health (Tanaka et al.
+1988).
+However, nivalenol (NIV), another type-B trichothecene, has been found in commercial wheat samples from Brazil (Del Ponte et al.
+2015) and also in inoculations of wheat under controlled conditions (Mendes et al.
+2018; Nicolli et al.
+2015).
+A molecular survey conducted during the last decade showed that *Fusarium graminearum* (O’Donnell et al.
+2004), formerly lineage 7 (O’Donnell et al.
+2000), is the main (>85% frequency) *F.
+graminearum* species complex (FGSC) member causing FHB of wheat in Brazil (Del Ponte et al.
+2015).
+The **F.
+graminearum** strains found in Brazilian wheat and barley analyzed so far are exclusively of the 15-acetyl(A)DON genotype, a mainly DON producer (Astolfi et al.
+2012; Castañares et al.
+2016; Del Ponte et al.
+2015).
+*Fusarium meridionale* is the second most dominant FHB pathogen (>15% overall) and an important regional contributor of FHB in the major wheat-producing region of Brazil, Paraná State, where its frequency increases to ~30% in wheat spikes and grain (Del Ponte et al.
+2015).
+This species is the main cause (> 96% frequency) of Gibberella ear rot and stalk rot as well the most frequent species surviving on maize stubbles, a large reservoir of inoculum for FHB epidemics (Del Ponte et al.
+2015; Kuhnem et al.
+2015).
+Other FGSC members that occur at smaller frequency in wheat include *F.
+cortaderiae* and *F.
+austroamericanum* that possesses either 3ADON or NIV trichothecene genotype (Del Ponte et al.
+2015).
+Nivalenol, as well as 3-ADON strains within FGSC members are actually much less common in wheat grown in the Americas, especially in Brazil (Del Ponte et al.
+2015), but there are concerns about their ability to produce different or greater amount of trichothecenes (Mendes et al.
+2018; Umpiérrez-Failache et al.
+2013) and potential shifts in climate that may favor build up of inoculum of the currently minor species (Juroszek et al.
+2013; Vaughan et al.
+2016).
+In the state of Louisiana, U.S., NIV-type populations of *F.
+graminearum*, which were found at high frequency (79%) among isolates from small-grain-growing regions, accumulated four times less trichothecene toxin than DON types on inoculated wheat (Gale et al.
+2011).
+The 3-ADON strains of *F.
+graminearum*, which appeared to emerge and spread into some regions of North America, were able to produce higher quantities of total trichothecenes and exhibit higher fecundity and growth rates than 15-ADON strains (Ward et al.
+2008).
+More recently, the aggressiveness of *F.
+graminearum* strains of the 15ADON or 3-ADON genotypes from New York towards a susceptible wheat cultivar did not differ, thus suggesting no apparent advantage of 3-ADON compared to 15-ADON genotype, although the former produced significantly greater amount of trichothecenes (Spolti et al.
+2014a).
 
-# Introduction
+In the south of Brazil, besides the dominant *F.
+graminearum* of the 15-ADON chemotype population, 3-ADON strains were found only in *F.
+cortaderiae* and *F.
+austroamericanum* isolates, while the NIV genotype was found in *F.
+cortaderiae*, *F.
+meridionale* and *F.
+asiaticum* (Del Ponte et al.
+2015).
+The 3-ADON strains from wheat were found at higher elevation regions with cooler winter and spring seasons, which suggested that climate could also shape species composition, besides cereal hosts (Del Ponte et al.
+2015).
+A recent study comparing 15-ADON and NIV strains of *F.
+graminearum* from China showed that the former grew faster at both 20 °C and 25 °C, produced higher amount of perithecia, and were more aggressive towards a wheat cultivar (Liu et al.
+2017).
+These findings of enhanced aggressiveness of the 15-ADON agree with studies in Brazil, although the NIV strains were of a different species (*F.
+meridionale*) (Spolti et al.
+2012a; Spolti et al.
+2013).
+The toxigenic potential of *F.
+graminearum* also appeared to be higher than F.
+meridionale because much greater amount of DON was produced compared to NIV produced by the latter (Nicolli et al.
+2015).
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Ut vehicula risus nec libero imperdiet fermentum.
-Maecenas convallis libero condimentum finibus euismod.
-Nunc lacinia, leo et sagittis posuere, augue sapien molestie lacus, ut imperdiet velit sem et nunc.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Aliquam euismod lorem vitae consequat congue.
-Donec bibendum, nulla et venenatis tincidunt, nulla quam hendrerit lorem, nec semper nunc tortor sit amet dui.
-Maecenas sit amet laoreet sem.
+Empirical knowledge of some components of the life cycle of other FGSC members than *F.
+graminearum* and F.
+meridionale in Brazil, such as the saprophytic ability, and whether this is influenced by the cereal-based substrate and temperature, which may provide competitive advantages for a species to build up inoculum for further spread are not available.
+In this study, a sample of 30 strains representative of three trichothecene genotypes among four FGSC members associated with FHB of wheat in southern Brazil was selected to investigate on the fitness-related traits related to both the saprophytic and pathogenic phase.
 
-Phasellus mattis mi et erat tristique, sit amet laoreet diam pulvinar.
-Sed rhoncus vulputate nibh, nec convallis metus viverra ultricies.
-Vivamus cursus feugiat ante, eget tincidunt dui tempus ac.
-Suspendisse a tempus leo.
-Curabitur varius neque finibus tincidunt aliquam.
-Orci varius natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
-Mauris eu ligula faucibus, ultricies risus vel, convallis mi.
-Aenean convallis vitae orci nec sollicitudin.
-Mauris mollis justo id vehicula elementum.
-Duis et neque eleifend, auctor ligula id, finibus ligula.
-Aliquam sagittis mauris a nisl consequat, eu mattis ante ultricies.
-Sed porttitor consequat placerat.
-Integer aliquam turpis nec tellus condimentum scelerisque.
-Nulla maximus metus ligula, sit amet sodales magna pellentesque nec.
-Cras accumsan pretium lectus a convallis.
+## Material and Methods
 
-Aliquam erat volutpat.
-Cras commodo eros arcu, ac pellentesque augue gravida at.
-Nulla feugiat dui nec ligula aliquam, ut lacinia ex facilisis.
-Nunc ut metus pulvinar, pulvinar sem vel, maximus lectus.
-Nulla commodo lorem vitae porttitor tempus.
-Nullam sed nibh sed eros aliquet congue.
-Proin sollicitudin vestibulum ex et tincidunt.
+## Isolates and experiments
 
-Nulla rutrum leo nec feugiat fringilla.
-Suspendisse sed hendrerit orci.
-Fusce quis lobortis nibh, eget luctus mi.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Nulla nec purus a lectus semper congue sit amet quis ipsum.
-Mauris accumsan dapibus auctor.
-Curabitur rhoncus massa eu felis ultricies mattis vel vel sem.
+The thirty isolates selected for this study were originated from surveys of symptomatic wheat spikes in commercial fields during four seasons (2007/08 to 2011/12) in southern Brazil.
+These were accurately identified to species and trichothecenes genotypes using PCR-based and multilocus-genotyping (MLGT) assays (Astolfi et al.
+2012; Del Ponte et al.
+2015).
+The strains were deposited and are available upon request at a mycological collection (CML - “Coleção Micológica de Lavras”).
+Table 1 shows detailed information for these strains and indicates in which experiments they were used.
+A total of five different experiments were conducted to assess saprophytic (sexual and asexual reproduction), pathogenic (severity towards two cultivars), and sensitivity to a triazole fungicide (tebuconazole).
 
-Aliquam sit amet lacinia massa, nec convallis ex.
-Nullam porttitor tincidunt mauris, elementum pellentesque velit tristique nec.
-Nunc a sollicitudin ipsum.
-Phasellus felis turpis, pretium vel scelerisque eget, ultricies id ex.
-Mauris in ullamcorper felis.
-Quisque nec malesuada urna.
-Integer pellentesque, magna vitae imperdiet ullamcorper, augue lacus blandit tortor, quis vulputate dolor elit non elit.
-Nullam euismod orci et rhoncus fringilla.
-Quisque vitae mauris sed eros sagittis lacinia.
-Vestibulum ipsum lorem, pretium a orci et, sagittis ultrices mi.
-Ut id eleifend libero.
-Fusce est ipsum, finibus id quam id, auctor tincidunt tortor.
-Suspendisse accumsan sit amet sapien at rutrum.
+### Perithecial production
 
-# Material and Methods
+Perithecial production, as indicative of sexual fertility, was evaluated using kernel-based substrates composed of a distinct cereal host: wheat, rice or maize.
+The method was adapted from Chen & Zhou (2009).
+Briefly, 20g of grain was soaked in 10 mL distilled water and autoclaved for 20 min at 127 ° C daily during consecutive days.
+Afterwards, kernels were inoculated with three discs (6 mm diameter) of mycelia and kept in the dark for 21 days at 25 °C.
+Then, the colonized grains were transferred to a bench in a greenhouse with controlled temperature (25 ±3 °C) where they remained for seven days for drying.
+After drying, they were placed in plastic box (11 x 11 x 3.5 cm) containing sterile sand layer (2 cm) and three sheets of filter paper on top of the sand.
+The filter paper was moistened daily with sterile distilled water.
+After 21 days, 50 grains were randomly selected from each substrate-isolate combination and individually scored visually using a diagrammatic key developed for this study and which contains four scores representing a percent range for the grain surface covered with perithecia (Supplementary Fig.
+1).
+The frequency of the scores for each species-genotype combination (five groups) was normalized to a perithecial production index (PPI, %).
 
-## item1
+### Mycelial growth
 
-Cras sollicitudin massa et felis rhoncus, scelerisque feugiat metus rutrum.
-Aenean blandit augue id sapien mattis consectetur.
-Nulla sollicitudin lectus nec massa iaculis, sed aliquet ante consequat.
-Donec ut vehicula felis.
-Etiam placerat convallis efficitur.
-Mauris id lacinia est.
-Vivamus vitae diam metus.
-Aliquam eget massa at ante ornare ultrices id nec tortor.
-Curabitur quam odio, tincidunt sit amet ipsum id, hendrerit viverra erat.
-Integer lacinia eros nec metus ultrices, vitae semper sapien rhoncus.
-Nulla id tempor ante.
-Nullam sem turpis, sagittis nec risus sed, imperdiet scelerisque lacus.
-Vivamus tincidunt felis a risus iaculis lacinia.
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
+Isolates grown on PDA (20 mL per dish) for seven days at 25 °C in continuous darkness provided mycelial discs (6 mm diameter) which were individually placed in the center of plastic Petri dishes (90 mm diameter).
+The cultures were incubated in growth chambers at two regimes of sub-optimal constant temperature (15 °C or 30 °C) under continuous darkness for five days when the diameter of the colony was measured from two perpendicular directions.
+The mycelial growth rate (mm/day) was calculated after subtracting the mycelium plug size (Spolti et al.
+2012b).
+Three replicates (plates) were used per treatment.
 
-## item2
-Vivamus ullamcorper justo sapien, vitae efficitur dui laoreet vitae.
-Etiam tristique, nibh vitae varius mattis, turpis leo elementum eros, quis dictum dolor enim a nunc.
-Etiam cursus velit eget nisl gravida, quis faucibus justo congue.
-Nunc vestibulum venenatis felis.
-Nullam sit amet nibh erat.
-Nulla vel ex viverra, fringilla ex eget, luctus enim.
-Nulla lacinia nisi risus, sit amet laoreet purus mattis quis.
-Aliquam pellentesque dolor ac metus blandit iaculis.
-Nulla mollis massa magna, non auctor nibh tempor vitae.
-Quisque mollis vel dolor eu pretium.
+### Sporulation and germination
 
-## item3
+The asexual fecundity of the strains was evaluated based on the macroconidia production in culture medium.
+Isolates were grown on SNA media for seven days under a 12/12 h light/dark cycle at 25 ° C (Leslie and Summerell, 2006).
+After this period, three discs of 6 mm in diameter were removed from the edges of the developing colonies and immersed in 5 mL of sterile distilled water containing 0.1 mL Tween 0.001% in a test tube and shaken for 20 s.
+The spore numbers were counted and the sporulation expressed as number of spores/mL.
+Spore germination was observed for a subsample of 20 spores (three replicated slides) after six hours of exposure to water and results were expressed as percent of germinating spores.
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Proin in blandit est.
-Fusce quis hendrerit libero, id auctor quam.
-Duis et vestibulum mauris.
-Maecenas sagittis condimentum laoreet.
-Donec vitae lectus eu erat hendrerit facilisis sed eget sem.
-Nam interdum libero ac ex cursus suscipit.
+### Pathogenicity
 
-# Results
+The pathogenicity of the strains was evaluated using a susceptible (BRS 194) and a moderately resistant (BRS Guamirim) cultivar.
+Seeds were sown in plastic pots (3 L) containing carbonized rice husk substrate (Tropstrato HT vegetables) fertilized with macronutrients (NPK 4-14-8).
+After seedling emergence, they were thinned to obtain 10 plants per pot.
+Inoculations were performed during the milk stage of grain development of the main tillers of the plants using the central-spikelet method (Engle et al.
+2003).
+The mid-position spikelet (sixth from top to bottom) was marked with a pen and 20 µL of a spore suspension (105 conidia/mL) was gently applied inside the floret with the aid of a micropipette (Spolti et al.
+2012b).
+After inoculation, the plants were kept at constant temperature (25 °C) to induce infection.
+The proportion of diseased spikelets in the wheat head was evaluated every three to five days during three weeks after inoculation.
+The experiment was repeated once in time and the data combined for analysis.
+The area under the diseased progress curve (AUDPC) was calculated and used to evaluate the effect of the trichothecene genotypes.
 
-## item1
+### Sensitivity to tebuconazole
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Proin in blandit est.
-Fusce quis hendrerit libero, id auctor quam.
-Duis et vestibulum mauris.
-Maecenas sagittis condimentum laoreet.
-Donec vitae lectus eu erat hendrerit facilisis sed eget sem.
-Nam interdum libero ac ex cursus suscipit.
+Tebuconazole sensitivity was quantified by measuring mycelial growth on PDA media (potato dextrose agar) amended with increasing fungicide concentrations (Becher et al.
+2010; Spolti et al.
+2014b).
+Stock solution (100 μg active ingredient [a.i.]/mL) was adjusted by dilution of commercial formulation (Folicur 200 EC, BAYER S.A., 20% a.i.) in distilled.
+Stock solution was added to cooled PDA to obtain the following tested concentrations of 0.0; 0.5; 1.0; 2.0, 4.0 and 8.0 μg a.i./mL.
+One mycelia agar disc (6 mm in diameter) obtained from the edge of a 7-day-old culture was placed at the central position of a Petri dish (90 mm) containing 15 mL of PDA amended with the fungicide at each concentration.
+After four days of incubation at 25°C in darkness, radial growth was measured in two perpendicular directions and the disc diameter was subtracted.
+For each repeat of the experiment, two replicates were used.
+The effective fungicide concentration to reduce 50% of mycelial growth (EC50) was determined using a linear regression model as described elsewhere (Becher et al.
+2010; Spolti et al.
+2014b).
 
-## item2
+### Statistical analyses
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Proin in blandit est.
-Fusce quis hendrerit libero, id auctor quam.
-Duis et vestibulum mauris.
-Maecenas sagittis condimentum laoreet.
-Donec vitae lectus eu erat hendrerit facilisis sed eget sem.
-Nam interdum libero ac ex cursus suscipit.
+For the data from the experiments to evaluate saprophytic growth (perithecia formation, mycelial growth and sporulation-germination) and pathogenicity, a multilevel linear mixed model was used to test the effects of the interaction, whenever present (e.g.
+temperature x genotype or species or species-genotype combination), or the single effect of a factor on the response variable at 0.05 significance level.
+In the model, replication for the isolates and isolates within each species-genotype or trichothecene genotype, depending on the experiment, were treated as random effects.
+The lsmeans and respective confidence intervals were obtained (using emmeans package of R) using kenward rogers df approximation.
+A Tukey method was used for comparing the means (alpha = 0.05) whenever the effect was deemed significant.
+Finally, the experiment to obtain EC50 values was repeated once in time and the data were combined for analysis.
+The means and confidence intervals were calculated.
 
-## item3
+## Results
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Proin in blandit est.
-Fusce quis hendrerit libero, id auctor quam.
-Duis et vestibulum mauris.
-Maecenas sagittis condimentum laoreet.
-Donec vitae lectus eu erat hendrerit facilisis sed eget sem.
-Nam interdum libero ac ex cursus suscipit.
+### Saprophytic fitness 
+The proportion of grains with scores >1 (or >10% kernel area covered with perithecia) was apparently higher for maize and rice kernels substrates compared with wheat and for the 15-ADON isolates compared with NIV and 3-ADON (Figure 1A).
+All interactions tested in the model (species, genotypes or species-genotype combined versus substrates) significantly affected (P <0.001) perithecial production index (PPI), which was highest (>50%) for 15-ADON *F.
+graminearum* strains inoculated on rice and maize, and lowest for inoculations on wheat kernels (Figure 1B).
+There was a general trend of decreasing PPI in the wheat kernel substrate compared with the rice and maize substrates, as well as in the 3-ADON strains compared with the 15-ADON strains.
+For the two NIV-producing species, PPI was higher in F.
+meridionale strains than in F.
+cortaderiae strains inoculated in rice and maize, but not in wheat.
+There was no difference (*P* > 0.05) in PPI between the two 3-ADON species, regardless of the substrate.
 
-# Discussion
+As expected, mean values of mycelial growth rate (MGR) was 84% higher at 30 °C (0.7 mm/day) than at 15 °C (1.2 mm/day) regardless of the trichothecene genotype.
+Within each of the temperatures, there were no differences in MGR means among the trichothecene genotypes (*P* > 0.20) as well as between species within trichothecene genotypes (Figure 2A).
+The variation in MGR was apparently higher within 15-ADON strains compared with the others, especially for the high temperature.
 
-Sed vitae neque diam.
-Donec facilisis nisi nec mollis finibus.
-Nullam fermentum, nisl luctus vehicula vestibulum, risus eros molestie enim, eget auctor nunc libero sit amet nulla.
-Suspendisse ac consectetur sapien.
-Sed lacinia, neque lacinia blandit congue, sapien erat malesuada augue, ac vestibulum quam metus id odio.
-Morbi metus magna, tempor quis enim id, commodo eleifend risus.
-Sed quis mollis lectus, sed imperdiet tortor.
-Pellentesque eget condimentum velit, a imperdiet ante.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae;
+All strains yielded macroconidia and there were differences in conidia production among the genotypes (P = 0.057), which were more evident when species-genotypes were compared (P < 0.001).
+The most prolific species was *F.
+meridionale* (2.18 x 103 spores/mL), which differed significantly from the other species (0.58 to 0.9 x 103 spores/mL), but not from *F.
+graminearum* (1.50 x 103 spores/mL) (Fig 2B).
 
-Vivamus ullamcorper justo sapien, vitae efficitur dui laoreet vitae.
-Etiam tristique, nibh vitae varius mattis, turpis leo elementum eros, quis dictum dolor enim a nunc.
-Etiam cursus velit eget nisl gravida, quis faucibus justo congue.
-Nunc vestibulum venenatis felis.
-Nullam sit amet nibh erat.
-Nulla vel ex viverra, fringilla ex eget, luctus enim.
-Nulla lacinia nisi risus, sit amet laoreet purus mattis quis.
-Aliquam pellentesque dolor ac metus blandit iaculis.
-Nulla mollis massa magna, non auctor nibh tempor vitae.
-Quisque mollis vel dolor eu pretium.
+The germination rate ranged from 0 to 100% across the strains.
+The greatest variability (80% range) was found for the *F.
+graminearum* strains, followed by F.
+austroamericanum for which the two strains differed by >70%.
+Therefore, the effect of genotype was not significant (P = 0.17), but when species-genotype was compared, the mean germination rate of F.
+meridionale (10.5%) differed significantly from *F.
+graminearum* (50%) and F.
+cortaderiae NIV (52.8%).
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Proin in blandit est.
-Fusce quis hendrerit libero, id auctor quam.
-Duis et vestibulum mauris.
-Maecenas sagittis condimentum laoreet.
-Donec vitae lectus eu erat hendrerit facilisis sed eget sem.
-Nam interdum libero ac ex cursus suscipit.
+### Pathogenicity
 
-Integer efficitur viverra elementum.
-Quisque commodo blandit blandit.
-Donec eget ex et ante iaculis porta.
-In sem augue, lacinia ut venenatis ullamcorper, volutpat quis tortor.
-Proin molestie ultrices purus ac consectetur.
-Duis ultrices, nunc vel euismod finibus, dui nunc elementum mauris, cursus vestibulum lorem magna in dui.
-Vivamus euismod consequat dolor.
+FHB severity (proportion of symptomatic spikelets) increased over time and ranged from 10% to 40% in the two cultivars.
+Regardless of the species, disease progress appeared slower during the first ten days after inoculation (<10%) for the moderately resistant cultivar (BRS Guamirim) compared to the susceptible (BRS 194), for which the progress peaked and stabilized at 15% to 35% after 10 days (Figure 3).
+No formal test was used to compare these two cultivars because they were evaluated in different experiments.
+Results of the mixed model to test the effect of genotype in the mean area under the severity curve showed the 15-ADON *F.
+graminearum* (432.1) was significantly higher (P < 0.05) than the 3-ADON- (331.03) and NIV-producing (319.6) species when inoculated in the moderately resistant cultivar, BRS Guamirim.
+On the other hand, the area under the severity curve in the BRS 194 inoculated with the 3-ADON genotypes (259.8, F.
+austroamericanum) was significantly higher than the NIV-producing strains (179.4); both not differing from the *F.
+graminearum* 15-ADON (204.38).
 
-Cras id sem augue.
-Nullam tincidunt vulputate arcu in dignissim.
-Vivamus gravida feugiat turpis at cursus.
-Donec euismod sit amet orci vel faucibus.
-In convallis rutrum leo eget elementum.
-Fusce mattis lacus quis risus volutpat, vel imperdiet ex dapibus.
-Quisque imperdiet est ac eros dapibus, dapibus aliquam justo porttitor.
+### Tebuconazole sensitivity
 
-# Acknowledgements
+The estimated EC50 values ranged from 0.007 to 1.42 µg/mL across the strains.
+The highest mean EC50 value was found for one 15-ADON strain (1.42 µg/mL), while for the others the values were < 0.5 µg/mL.
+Differences among the genotypes were not apparent across the strains and the limited number of samples prevented formal comparison.
 
- Fusce quis hendrerit libero, id auctor quam.
-Duis et vestibulum mauris.
-Maecenas sagittis condimentum laoreet.
-Donec vitae lectus eu erat hendrerit facilisis sed eget sem.
-Nam interdum libero ac ex cursus suscipit.
+## Discussion
 
-# References
+In this study we investigated for the first time several traits related to different stages of the saprophytic and pathogenic phases of the FGSC life cycle using a sample of strains representative of four members (and three trichothecene chemotypes) associated with FHB in Brazil.
+At least for some components evaluated, results are supportive of the hypothesis that *F.
+graminearum* is more fit than the other species, which may help to explain its dominance as FHB pathogen of wheat in Brazil (Astolfi et al.
+2012; Del Ponte et al.
+2015; Scoz et al.
+2009).
+
+Our most striking result favoring *F.
+graminearum* was sexual fertility.
+These heterothallic fungi are able to overwinter in crop residues where they produce perithecia and ascospores which are able to disperse long-distance and infect wheat starting at flowering (Gilbert & Fernando, 2004; Khonga & Sutton, 2009; Pereyra & Dill-Macky, 2004).
+Thus, the perithecia production capacity should play an important role in the ecology and epidemiology (Leplat et al., 2013; McMullen et al, 2012).
+So far, few studies have characterized and compared FGSC members or chemotypes within a same species, from a same region/crop, with regards to perithecia and ascospore production (Lee et al.
+2012; Liu et al.
+2017; Spolti et al.
+2014a).
+Similarly to others (Lee et al.
+2009; Liu et al.
+2017), we inferred on fertility based on the amount of perithecia produced on grain from three main cereal substrates.
+We found that *F.
+graminearum*, the only member that possess the 15-ADON trichothecene genotype in Brazil, produced higher amount of perithecia, followed by the NIV-producing *F.
+meridionale*, and then by the other two species with the 3-ADON or NIV chemotype.
+Interestingly, this ranking strongly agrees with that of the species frequency found in wheat and barley grain in the region (Astolfi et al., 2011; Astolfi et al., 2012; Del Ponte et al., 2015).
+
+Previously, the ability of a dominant kernel-born FGSC member to produce significantly large amounts of perithecia in the original host was observed by Lee et al.
+(2009), who isolated 87% of *F.
+asiaticum* (main species found in kernels) strains from perithecia developing on rice plants after physiological maturity, followed by *F.
+graminearum* (13%).
+Recently, Liu et al (2017) compared perithecia production by *F.
+asiaticum* (NIV and 3-ADON) and *F.
+graminearum* (15-ADON) from China and found that the latter produced the highest amount of perithecia, which were visible two to three days earlier than the NIV and 3-ADON strains.
+In that Chinese region, *F.
+graminearum* 15-ADON is also dominant in wheat kernels.
+In New York, a comparison between 15-ADON and 3-ADON strains of *F.
+graminearum* from wheat showed no difference in saprophytic and pathogenic fitness towards wheat (Spolti et al.
+2014a).
+
+Previously, we found that perithecium-borne *F.
+meridionale* was highly prevalent in maize stubbles over the surface of two out of three wheat fields, and *F.
+graminearum* absent or at very low frequency in stubbles of all three fields (Del Ponte et al., 2015).
+We then hypothesized that the former was more fit towards maize, which was further confirmed more recently, as *F.
+meridionale* was largely dominant on stubbles, stalks and kernels of maize from Brazil (Kuhnem et al.
+2016).
+In the wheat study, although *F.
+meridionale* was dominant in perithecia from maize stubbles on the surface of wheat fields, it was the least frequent in wheat kernels from the same field, which were largely infected by *F.
+graminearum* (Del Ponte et al.
+2015).
+Our results using three cereal-based substrates suggest that the dominance is not due to enhanced capacity of F.
+meridionale to produce perithecia when compared to *F.
+graminearum*, but to its ability to primarily cause stalk rot.
+Further studies are underway to investigate whether *F.
+meridionale* is able to outcompete *F.
+graminearum* in both single and mixed inoculations of maize stalks.
+
+A few other questions remain.
+It would be instructive to further examine ascospore production and the aerobiology of these species in both controlled and field environment.
+In our previous field study, *F.
+meridionale* was the least frequent (<30%) species collected in selective media spore traps during wheat flowering (Del Ponte et al.
+2015).
+Ideally, new studies should include more accurate or less-biased methods, such as molecular identification, given that during isolations, a fast-growing species may outcompete others.
+In addition, it would be important to monitor ascorporic inoculum over time and space when both species are present in similar frequency, similar to another study that used spore sampler placed near maize stalk residues bearing perithecia with mature ascospores (Manstretta & Rossi 2015).
+It is well know that FGSC are capable of long distance dispersal (Prussin et al.
+2014) and that airborne populations above wheat canopies may be composed of well-mixed inoculum originated under the effect of turbulence and gravitational settling of inoculum from both local and external sources (Del Ponte et al.
+2003; Schmale et al.
+2006).
+The latter hypothesis has been suggested based on both aerobiological and spatial patterns studies in Brazil (Del Ponte et al.
+2005; Spolti et al.
+2015).
+
+ There was a large variation among the isolates from a same species-chemotype with regards to the response to the two sub-optimal temperatures tested in our study, thus preventing us from detecting differences in mycelial growth rate at the species or chemotype level.
+Previous studies have shown a large sensitivity of *F.
+graminearum* strains to temperature increase, which was indicative of potential for thermal adaptation (Backhouse et al.
+2014; Spolti et al.
+2014a; Zhan & McDonald, 2011) and saprophytic adaptability (Tunali et al.
+2012).
+In our previous work evaluating mycelial growth rate of only two *F.
+graminearum* and two F.
+meridionale strains, growing at four increasing temperatures (10 to 30 oC), significant difference between species was found only at 25 oC and 30 oC, with *F.
+graminearum* growing faster than *F.
+meridionale* (Spolti et al.
+2012a).
+Using a sample as low as eight strains, the maximum difference among strains of *F.
+graminearum* and *F.
+meridionale* found in this study was ~10 mm/day, which suggests caution when inferring species differences based on small sample.
+
+Differences in macroconidia production in culture among species may be suggestive of advantages for further spread by assexual means (Leplat et al.
+2013).
+We found that the variation in this trait was very large among isolates of the same species, especially *F.
+graminearum* and *F.
+meridionale*, the most represented species, which prevented us from detecting differences as found in our previous studies using a different set of isolates and smaller sample (Spolti et al.
+2012a; Spolti & Del Ponte 2013).
+Recently, Liu et al.
+(2017) reported significantly greater macroconidia production by 20 *F.
+graminearum* 15-ADON strains than 20 *F.
+asiaticum* NIV strains, but they both did not differ from the F.
+asiaticum 3-ADON strains.
+The germination rate also varied considerably among strains of the same species, which makes largely unclear whether macroconidia production is a reliable variable to compare the asexual reproduction fitness among members of the complex given the inconsistencies and high variability.
+Moreover, in natural epidemics caused by FGSC, the importance of macroconidia inoculum is not clear, although studies highlighted the potential to dispersal and further infection (Paul et al.
+2007).
+
+We found that the species with the DON chemotype were generally more aggressive (higher area under severity curve) than the NIV ones, which is in agreement with our previous studies (Mendes et al.
+2018; Spolti et al.
+2012a).
+We could not compare statistically the effect of cultivar because the experiments were conducted at different times, and with different total time duration of the evaluations.
+Nevertheless, the area under the curve was in general higher for BRS 194, the susceptible one, compared with BRS Guamirim as shown earlier (Mendes et al.
+2018; Spolti et al.
+2012a).
+For the first time, we compared the two acetylates of DON within the Brazilian strains; while 15-ADON (*F.
+graminearum*) was the most aggressive in a susceptible cultivar, the 3-ADON (*F.
+austroamericanum*) was the most aggressive in the moderately resistant cultivar.
+Previously, studies comparing strains of the these two acetylates of DON showed differences in the rate of colonization, with 15-ADON being more aggressive than NIV, only in the less susceptible (BRS Guamirim), and not in BRS 194  (Spolti et al.
+2012a).
+
+We used the area under the curve of severity and not the final severity (last assessment) (Spolti et al.
+2012a), or count of diseased spikelets at individual times (Liu et al.
+2017) to compare the strains.
+Although these measures may be correlated, the area under the curve is more appropriate to capture differences due to variations in the daily progress (e.g.
+different curve shapes may lead to a same final severity).
+It would be important to standardize the methodology for assessing FHB pathogen aggressiveness, especially the type of variables and analysis when using central spikelet inoculation in order to compare data from multiple studies.
+These would include: inoculation timing, spore concentration, spikelet position, disease assessment, area under curve of final severity or disease count, etc.
+A previous study has shown that, in some cases, the bleaching above the inoculation point for some cultivars not due to fungal spread but to death of vascular tissue, which is hard to discern visually (Brown et al.
+2010; Malbrán et al.
+2014).
+During our observations, we noticed that the total number of spikelets varied among the spikes (data not shown).
+Therefore, the percent severity may vary slightly when the same number of diseased spikelets is observed in heads of different sizes.
+The advantages and technical (statistical) issues of using count or proportion data is a topic that deserves further investigation.
+
+We provide data on the tebuconazole sensitivity for five FGSC-chemotype found in southern Brazil, but we acknowledge the limitations of our sample sizes.
+Thus far, few studies evaluated the overall fungicide sensitivity of at least two FGSC members from wheat in Brazil and used either small or more representative sampling (>50 strains) (Spolti et al.
+2012a; Spolti et al.
+2012b).
+The range of EC50 values found in our study varied not only among, but also within isolates of a same species, but they were within the range reported for *F.
+graminearum* isolates from the same region (Spolti et al.
+2012a).
+Previously, differences in tebuconazole sensitivity between 3-ADON and 15-ADON strains within *F.
+graminearum* from NY State were not found when using large sample (25 strains each) (Spolti et al.
+2014b).
+Although our small sample, which is far from being representative of the species, *F.
+graminearum* was in general less sensitive to tebuconazole because of one less sensitive isolate, than the other less frequent species.
+This agrees specifically with a recent study from our group where a larger sample of 15-ADON (48 strains) and the other two genotypes (NIV + 3-ADON = 17 strains) from barley was used (Machado et al.
+2017).
+Conversely, in Uruguay, the NIV-producing strains within *F.
+asiaticum* and *F.
+cortaderiae* were significantly less sensitive than 15-ADON *F.
+graminearum* strains.
+Further molecular studies would be important to investigate the genetic basis of such differences in triazole sensitivity among the FGSC members.
+
+Our results contribute new knowledge to improve our understanding of the ecology and epidemiology of *F.
+graminearum* species complex members, which may be of value for improving models for assessing the risk or epidemics and mycotoxin production.
+These should include assessing whether there are species-specific environmental requirements for fungal infection, colonization and mycotoxin production.
+
+## Acknowledgements
+
+First author is grateful to the Programa de Pós-graduação em Fitotecnia (UFRGS) and CNPq - Conselho Nacional de desenvolvimento Científico e Tecnológico, for providing a graduate scholarship CNPq.
+
+## References
+
+Astolfi, P., Dos Santos, J., Gomes, L.B., Silva, C.N., Tessmann, D.J., Del Ponte, E.M.
+2011.
+Molecular survey of trichothecene genotypes of *Fusarium graminearum* species complex from barley in southern Brazil.
+Int J Food Microbiol.
+148:197-201.
+
+Astolfi, P., Reynoso, M.M., Ramirez, M.L., Chulze, S.N., Alves, T.C.A., Tessmann, D.J., Del Ponte, E.M.
+2012.
+Genetic population structure and trichothecene genotypes of *Fusarium graminearum* isolated from wheat in southern Brazil.
+Plant Pathol.
+61:289-295.
+
+Backhouse, D. 2014. Global distribution of *Fusarium graminearum, F. asiaticum* and *F. boothii* from wheat in relation to climate. European Journal of Plant Pathology 139: 161-173.
+
+Becher, R., Hettwer, U., Karlovsky, P., Deising, HB., Wirsel, SGR.
+2010.
+Adaptation of *Fusarium graminearum* to tebuconazole yielded descendants diverging for levels of fitness, fungicide resistance, virulence, and mycotoxin production.
+Phytopathology.
+100:444-453.
+
+Brown, N.
+A., Urban, M., van de Meene, A.
+M.
+L., Hammond-Kosack, K.
+E.
+2010.
+The infection biology of *Fusarium graminearum*: Defining the pathways of spikelet to spikelet colonisation in wheat ears.
+Fungal Biology 114: 555–571.
+
+Castañares, E., Dinolfo, M.I., Del Ponte, E.M., Pan D., Stenglein, S.A.
+2016.
+Species composition and genetic structure of *Fusarium graminearum* species complex populations affecting the main barley growing regions of South America.
+Plant Pathol.
+65:930-939.
+
+Chen, Y., Zhou, M. G. 2009. Characterization of *Fusarium graminearum* isolates resistant to both carbendazim and a new fungicide JS399-19. Phytopathology 99:441-446.
+
+Del Ponte, E.M., Fernandes, J.M.C., Bergstrom, G.C.
+2003.
+Fusarium Head Blight and deoxynivalenol accumulation in wheat inoculated at developmental stages from flowering through grain maturation.
+Anais, Fusarium Head Blight Forum, St.
+Paul, MN, 129-132.
+
+Del Ponte, E.M., Fernandes, J.M.C., Pavan, W. 2005. A risk infection simulation model for fusarium head blight of wheat. Fitopatol. Bras. 30:634-642.
+
+Del Ponte, E.M., Spolti, P., Ward, T.J., Gomes, L.B., Nicolli, C.P., Kuhnem, P.R., Silva, C.N., Tessmann, D.J.
+2015.
+Regional and field-specific factors affect the composition of Fusarium head blight pathogens in subtropical no-till wheat agroecosystem of Brazil.
+Phytopathology 105:246-254.
+
+Engle, J. S., Madden, V. L., Lipps, P. E. 2003. Evaluation of inoculation methods to determine resistance reactions of wheat to *Fusarium graminearum*. Plant Dis. 87:1530-1535.
+
+Gale, L.R., Harrison, S.A., Ward, T.J., O’Donnell, K., Milus, E.A., Gale, S.W., Kistler, H.C.
+2011.
+Nivalenol-Type Populations of *Fusarium graminearum* and *F.
+asiaticum* are prevalent on wheat in Southern Louisiana.
+Phytopathology 101:124-134.
+
+Gilbert, J., Fernando, W.D.G. 2004. Epidemiology and biological control of *Gibberella zeae⁄Fusarium graminearum*. Can J Plant Pathol. 26:464–72.
+
+Juroszek, P., Von Tiedemann, A. 2013. Climate change and potential future risks through wheat diseases: a review. Eur J Plant Pathol. 136:21-33.
+
+Kuhnem, P.R., Del Ponte, E.M., Dong, Y., Bergstrom, G.C.
+2015.
+*Fusarium graminearum* isolates from wheat and maize in New York show similar range of aggressiveness and toxigenicity in cross-species pathogenicity tests.
+Phytopathology 105:441-448.
+
+Kuhnem, P.R., Ward, T.J., Silva, C.N., Spolti, P., Ciliato, M.L., Tessmann, D.J., Del Ponte, E.M.
+2016.
+Composition and toxigenic potential of the *Fusarium graminearum* species complex from maize ears, stalks and stubble in Brazil.
+Plant Pathol.
+65:1185-1191.
+
+Khonga, E.B., Sutton, J.C. 2009. Inoculum production and survival of *Gibberella zeae* in maize and wheat residues. Can J Plant Pathol. 10:232-239.
+
+Lee, J., Chang, I.Y., Kim, H., Yun, S.H., Leslie, J.F., Lee, Y.W. 2009. Genetic diversity and fitness of *Fusarium graminearum* populations from rice in Korea. Appl. Environ. Microbiol. 75:3289-3295.
+
+Lee, J., Kim, H., Jin Jeon, J., Kim, J.J., Zeller, K.A., Carter, L.L.A., Leslie, J.F.
+Leeb, Y.W.
+2012.
+Population Structure of and Mycotoxin Production by *Fusarium graminearum* from Maize in South Korea.
+Appl.
+Environ.
+Microbiol.
+78:2161-2167.
+
+Leplat, J., Friberg, H., Abid, M., Steinberg, C. 2013. Survival of *Fusarium graminearum* the causal agent of Fusarium head blight. A review. Agron. Sustain. Dev. 33:97-111.
+
+Leslie, J. F., Summerell, B. A. 2006. The *Fusarium* Laboratory Manual. Ames, Iowa: Blackwell Publishing, p.388.
+
+Liu, Y.Y., Sun, H.Y., Li, W., Xia, Y.L., Deng, Y.Y., Zhang, A.X.,Chen, H.G.
+2017.
+Fitness of three chemotypes of *Fusarium graminearum* species complex in major winter wheat-producing areas of China.
+PLoS One 17:0174040.
+
+Machado, F.M., Santana F.M., Lau D., Del Ponte E.M.
+2017.
+Quantitative review of the effects of triazole and benzimidazole fungicides on Fusarium Head Blight and wheat yield in Brazil.
+Plant Dis.
+101:1633-1641.
+
+Malbrán, L., Mourelos, C.A., Girotti, J.R., Balatti, P.A., Lori, G.A.
+2014.
+Toxigenic capacity and trichothecene production by *Fusarium graminearum* isolates from Argentina and their relationship with aggressiveness and fungal expansion in the wheat spike.
+Phytopathology 104:357-364.
+
+McMullen, M., Bergstrom, G., De Wolf, E., Dill-Macky, R., Hershman, D., Shaner, G., Sanford, D.V.
+2012.
+A unified effort to fight an enemy of wheat and barley: Fusarium head blight.
+Plant Dis.
+96:1712-1728.
+
+Mendes, G.R.L., Del Ponte, E.M., Feltrin, A.C., Badiale-Furlonge, E., Oliveira, A.C.
+2018.
+Common resistance to FHB caused by *Fusarium graminearum* and *Fusarium meridionale* in Brazilian wheat cultivars.
+Sci.
+Agric.
+In Press.
+
+Nicolli, C.P., Spolti, P., Tibola C.S., Fernandes J.M.C., Del Ponte, E.M.
+2015.
+Fusarium head blight and trichothecene production in wheat by *Fusarium graminearum* and *F.
+meridionale* applied alone or in mixture at post-flowering.
+Trop Plant Pathol.
+40:134-140.
+
+O'Donnell, K., Kistler, H.C., Tacke, B.K., Casper, H.H.
+2000.
+Gene genealogies reveal global phylogeographic structure and reproductive isolation among lineages of *Fusarium graminearum*, the fungus causing wheat scab.
+Proc.
+Natl.
+Acad.
+Sci.
+U.S.A.
+97:7905-79.
+
+O’Donnell, K., Ward, T.J., Geiser, D.M., Corby Kistler, H., Aoki, T.
+2004.
+Genealogical concordance between the mating type locus and seven other nuclear genes supports formal recognition of nine phylogenetically distinct species within the *Fusarium graminearum* clade.
+Fungal Genet Biol.
+41:600–623.
+
+Paul, P.A., Lipps, P.E., Hershman, D.E., McMullen, M.P., Draper, M.A., Madden, L.V.
+A.
+2007.
+Quantitative review of tebuconazole effect on Fusarium Head Blight and deoxynivalenol content in Wheat.
+Phytopathology 97:211-220.
+
+Pereyra, S.A., Dill-Macky, R., Sims, A.L. 2004. Survival and Inoculum Production of *Gibberella zeae* in Wheat Residue. Plant Dis. 88:724-730.
+
+Prussin II, A.J., Li, Q., Malla, R., Ross, S.D., Schmale III, D.G. 2014. Monitoring the long-distance transport of *Fusarium graminearum* from field-scale sources of inoculum. Plant Dis. 98:504-511.
+
+Schmale, D.G., III, Leslie, J.F., Zeller, K.A., Saleh, A.A., Shields, E.J., Bergstrom, G.C. 2006. Genetic structure of atmospheric populations of *Gibberella zeae*. Phytopathology 96:1021-1026.
+
+Scoz, L.B., Astolfi, P., Reartes, D.S, Schmale III, D.G., Moraes, M.G., Del Ponte, E.M.
+2009.
+Trichothecene mycotoxin genotypes of *Fusarium graminearum* sensu stricto and *Fusarium meridionale* in wheat from Southern Brazil.
+Plant Pathol.
+58:344-351.
+
+Spolti, P, Barros, N.C, Gomes, L.B, Santos, J, Del Ponte, E.M.
+2012a.
+Phenotypic and pathogenic traits of two species of the *Fusarium graminearum* complex possessing either 15-ADON or NIV genotype.
+Eur J of Plant Pathol.
+133:621-629.
+
+Spolti, P., Jorge, B.C., Del Ponte, E.M. 2012b. Sensitivity of *Fusarium graminearum* causing head blight of wheat in Brazil to tebuconazole and metconazole fungicides. Trop Plant Pathol. 37:419-423.
+
+Spolti, P., Del PontE, E.M.
+2013.
+Agressividade diferencial de espécies do complexo *Fusarium graminearum* em interação com o fungicida tebuconazole na redução do rendimento de trigo.
+Ciênc.
+Rural 43:1569-1575.
+
+Spolti, P., Del Ponte, E.M.D, Cummings, J.A., Dong, Y., Bergstrom, G.C.
+2014a.
+Fitness attributes of *Fusarium graminearum* isolates from wheat in New York possessing a 3-ADON or 15-ADON trichothecene genotype.
+Phytopathology 104:513-519.
+
+Spolti, P., Del Ponte, E.M., Dong, Y., Bergstrom, G.
+2014b.
+Triazole sensitivity in a contemporary population of Fusarium graminearum from New York wheat and competitiveness of a tebuconazole-resistant Isolate.
+Plant Dis.
+98:607-613.
+
+Spolti, P., Fernandes, J.M.C., BergstroM, G.C., Del Ponte, E.M.
+2015.
+Disease risk, spatial patterns, and incidence-severity relationships of Fusarium head blight in no-till spring wheat following maize or soybean.
+Plant Dis.
+99:1360-1366.
+
+Tanaka, T., Hasegawa, A., Yamamoto, S., Lee, U., Ugiura, Y., Ueno, Y.
+1988.
+Worldwide contamination of cereals by the *Fusarium* mycotoxins nivalenol, deoxynivalenol and zearalenone.
+1.
+Survey of 19 countries.
+J.
+Agric.
+Food Chem.
+36:979–983.
+
+Tunali, B., Obanor, F., Erginbas, G., Westecott, R.A., Nicol, J., Chakraborty, S. 2012. Fitness of three Fusarium pathogens of wheat. Microbiol Ecol. 81:596-609.
+
+Umpiérrez-Failache, M., Garmendia, G., Pereyra, S., Rodríguez-Haralambides, A., Ward, T.J., Vero, S.
+2013.
+Regional differences in species composition and toxigenic potential among Fusarium head blight isolates from Uruguay indicate a risk of nivalenol contamination in new wheat production areas.
+Intel J Food Microbiol.
+166:135-140.
+
+Vaughan, M., Backhouse, D., Del Ponte, E.M.
+2016.
+Climate change impacts on the ecology of *Fusarium graminearum* species complex and susceptibility of wheat to Fusarium head blight: a review.
+World Mycotoxin J.
+9:685-700.
+
+Ward, T.
+J., Clear, R.M., Rooney, A.P., O'Donnell, K., Gaba, D., Patrick, S., Starkey, D.E., Gilbert, J., Geiser, D.M., Nowicki, T.W.
+2008.
+An adaptive evolutionary shift in Fusarium head blight pathogen populations is driving the rapid spread of more toxigenic *Fusarium graminearum* in North America.
+Fungal Genet.
+Biol.
+45:473-484.
+
+Zhan, J., McDonald, B.A. 2011. Thermal adaptation in the fungal pathogen *Mycosphaerella graminicola*. Mol. Ecol. 20:1689-1701.
+
+Table 1.
+Information for a collection of 30 strains belonging to four phylogenetic species of the *Fusarium graminearum* species complex isolated from symptomatic wheat heads from fields in Rio Grande do Sul State, Brazil, which were used in five different experiments in this study
+
+Code | Location | 	Year  of isolation | 	Species |	Genotype |	Experiment
+
+3065	| Panambi	2009	| *F. graminearum* | 	15-ADON	| 2, 3, 4 
+3066	| Ijuí	2009 |	*F. graminearum*	| 15-ADON	| 1,4
+3067	Carazinho	2010	*F. graminearum*	15-ADON	1,2,3,4
+3071	Tapejara	2010	*F. graminearum*	15-ADON	1,2,3,4,5
+3069	Coxilha	2010	*F. graminearum*	15-ADON	1, 2, 4,5
+3070	Santa Bárbara do Sul	2011	*F. graminearum*	15-ADON	1, 2, 4
+3830	Não-Me-Toque	2011	*F. graminearum*	15-ADON	1,2,3
+3833	Não-Me-Toque	2011	*F. graminearum*	15-ADON	1
+3064	Cruz Alta	2007	*F. graminearum*	15-ADON	1,4,5
+3068	Ernestina	2007	*F. graminearum*	15-ADON	1,2,3,4,5
+3382	Caseiros	2010	F. austroamericanum	3-ADON	1,2,3,5
+3384	Barreto	2010	F. austroamericanum	3-ADON	1,2,3,4,5
+3379	Carazinho	2010	F. cortaderiae	3-ADON	1,2,3,4,5
+3845	Lagoa Vermelha	2010	F. cortaderiae	3-ADON	1,2,3,4
+3841	Lagoa Vermelha	2010	F. cortaderiae	3-ADON	1,2,3,4,5
+3831	Panambi	2009	F. cortaderiae	NIV	1,2,3,4,5
+3836	Coronel Barros	2009	F. cortaderiae	NIV	1,2,4
+3837	Coronel Barros	2009	F. cortaderiae	NIV	1,2,3,4,5
+3840	Mato Castelhano	2010	F. cortaderiae	NIV	1,2,4
+3834	Barreto	2010	F. cortaderiae	NIV	1,2,4
+3378	Vacaria	2010	F. cortaderiae	NIV	1,2,3,4,5
+3835	Ernestina	2007	F. cortaderiae	NIV	1,2,3,4
+3375	Santa Bárbara do Sul	2009	F. meridionale	NIV	1,2,4
+3843	Ernestina	2009	F. meridionale	NIV	1,2,3,4
+3344	Coxilha	2009	F. meridionale	NIV	1,2,3,4,5
+3381	Caseiros	2010	F. meridionale	NIV	1,2,3,4,5
+3839	Marau	2011	F. meridionale	NIV	1,2,3,4,5
+3377	Sertão	2011	F. meridionale	NIV	1,2,3,4
+3838	Nonoai	2007	F. meridionale	NIV	1,2,3,4
+3374	Nonoai	2007	F. meridionale	NIV	1,2,4
+
+a Culture collection code: CML = Coleção Micológica de Lavras, Universidade Federal de Lavras, Lavras, Minas Gerais, Brazil.
+b Trichothecene genotype identified by the Luminex method: 15-ADON = 15-acetyl-deoxynivalenol; 3-ADON = 3-acetyl-deoxynivalenol; NIV = nivalenol.
+c Experiments: 1 = Perithecial production; 2 = Mycelial growth; 3 = Sporulation and germination; 4 = Pathogenicity; 5 = Sensitivity to tebuconazole.
+
+Figure 1.
+Relative frequency of perithecial production scores based on a visual ordinal (0-3) scale for the proportion of grains in each score for the three different cereal based substrates (wheat, rice or maize kernels) (A) and the least square means and 95% confidence interval for the perithecial production index which is a normalization of the frequency of the scores (B).
+
+Figure 2.
+Mycelial growth and spore production by strains of four Fusarium graminearum species complex members possessing a nivalenol (NIV) or an acetylate form of deoxynivalenol (DON) genotype (3-ADON or 15-ADON) assessed based on mycelial growth at 15° C or 30° C (A), macroconidia production (B) and conidia germination (C).
+
+Figure 3.
+Mean (standard deviation) severity of Fusarium head blight (FHB) in a moderately resistant cultivar, BRS Guamirim (A), and a susceptible cultivar, BRS 194 (B), inoculated with strains of different Fusarium graminearum species complex members possessing a nivalenol (NIV) or an acetylate form of deoxynivalenol (DON) genotype (3-ADON or 15-ADON) found associated with FHB in Brazil.
+
+Figure 5.
+The effective concentration of tebuconazole that reduces 50% (EC50) of the mycelial growth of four Fusarium graminearum species complex members possessing a nivalenol (NIV) or an acetylate form of deoxynivalenol (DON) trichothecene genotype (3-ADON or 15-ADON) found associated with Fusarium head blight of wheat in Brazil.
+
+Supplemental material
+
+Figure 1S
+
+Diagrammatic key with four scores representing a percent range for the grain surface covered with perithecia, for rice, maize and wheat substrates.
+1 = 0% (no perithecia); 2 = 1 - 10% ; 3 = > 10% - 30%); 4 = > 30%.

### 2017-12-16 manuscript.Rmd
new style

@@
 A recent study comparing 15-ADON and NIV strains of *F.
-graminearum* from China showed that the former grew faster at both 20 °C and 25 °C, produced higher amount of perithecia, and were more aggressive towards a wheat cultivar (Liu et al.
+graminearum* from China showed that the former grew faster at both 20 <sup>o</sup>C and 25 <sup>o</sup>C, produced higher amount of perithecia, and were more aggressive towards a wheat cultivar (Liu et al.
 2017).
@@
 meridionale in Brazil, such as the saprophytic ability, and whether this is influenced by the cereal-based substrate and temperature, which may provide competitive advantages for a species to build up inoculum for further spread are not available.
-In this study, a sample of 30 strains representative of three trichothecene genotypes among four FGSC members associated with FHB of wheat in southern Brazil was selected to investigate on the fitness-related traits related to both the saprophytic and pathogenic phase.
+In this study, a sample of 30 strains representative of three trichothecene genotypes among four FGSC members associated with FHB of wheat in southern Brazil was selected to investigate the fitness-related traits related to both the saprophytic and pathogenic phase.
 
@@
 
+**Table 1.** Information for a collection of 30 strains belonging to four phylogenetic species of the Fusarium graminearum species complex isolated from symptomatic wheat heads from fields in Rio Grande do Sul State, Brazil, which were used in five different experiments in this study
+
+| CML Code<sup>a</sup> | Location             | Year of isolation | Species             | Genotype<sup>b</sup> | Experiment<sup>c</sup> |
+|------|----------------------|-------------------|---------------------|----------|------------|
+| 3065 | Panambi              | 2009              | F. graminearum      | 15-ADON  | 2, 3, 4    |
+| 3066 | Ijuí                 | 2009              | F. graminearum      | 15-ADON  | 1,4        |
+| 3067 | Carazinho            | 2010              | F. graminearum      | 15-ADON  | 1,2,3,4    |
+| 3071 | Tapejara             | 2010              | F. graminearum      | 15-ADON  | 1,2,3,4,5  |
+| 3069 | Coxilha              | 2010              | F. graminearum      | 15-ADON  | 1, 2, 4,5  |
+| 3070 | Santa Bárbara do Sul | 2011              | F. graminearum      | 15-ADON  | 1, 2, 4    |
+| 3830 | Não-Me-Toque         | 2011              | F. graminearum      | 15-ADON  | 1,2,3      |
+| 3833 | Não-Me-Toque         | 2011              | F. graminearum      | 15-ADON  | 1          |
+| 3064 | Cruz Alta            | 2007              | F. graminearum      | 15-ADON  | 1,4,5      |
+| 3068 | Ernestina            | 2007              | F. graminearum      | 15-ADON  | 1,2,3,4,5  |
+| 3382 | Caseiros             | 2010              | F. austroamericanum | 3-ADON   | 1,2,3,5    |
+| 3384 | Barreto              | 2010              | F. austroamericanum | 3-ADON   | 1,2,3,4,5  |
+| 3379 | Carazinho            | 2010              | F. cortaderiae      | 3-ADON   | 1,2,3,4,5  |
+| 3845 | Lagoa Vermelha       | 2010              | F. cortaderiae      | 3-ADON   | 1,2,3,4    |
+| 3841 | Lagoa Vermelha       | 2010              | F. cortaderiae      | 3-ADON   | 1,2,3,4,5  |
+| 3831 | Panambi              | 2009              | F. cortaderiae      | NIV      | 1,2,3,4,5  |
+| 3836 | Coronel Barros       | 2009              | F. cortaderiae      | NIV      | 1,2,4      |
+| 3837 | Coronel Barros       | 2009              | F. cortaderiae      | NIV      | 1,2,3,4,5  |
+| 3840 | Mato Castelhano      | 2010              | F. cortaderiae      | NIV      | 1,2,4      |
+| 3834 | Barreto              | 2010              | F. cortaderiae      | NIV      | 1,2,4      |
+| 3378 | Vacaria              | 2010              | F. cortaderiae      | NIV      | 1,2,3,4,5  |
+| 3835 | Ernestina            | 2007              | F. cortaderiae      | NIV      | 1,2,3,4    |
+| 3375 | Santa Bárbara do Sul | 2009              | F. meridionale      | NIV      | 1,2,4      |
+| 3843 | Ernestina            | 2009              | F. meridionale      | NIV      | 1,2,3,4    |
+| 3344 | Coxilha              | 2009              | F. meridionale      | NIV      | 1,2,3,4,5  |
+| 3381 | Caseiros             | 2010              | F. meridionale      | NIV      | 1,2,3,4,5  |
+| 3839 | Marau                | 2011              | F. meridionale      | NIV      | 1,2,3,4,5  |
+| 3377 | Sertão               | 2011              | F. meridionale      | NIV      | 1,2,3,4    |
+| 3838 | Nonoai               | 2007              | F. meridionale      | NIV      | 1,2,3,4    |
+| 3374 | Nonoai               | 2007              | F. meridionale      | NIV      | 1,2,4      |
+
+<sup>a</sup> Culture collection code: CML = Coleção Micológica de Lavras, Universidade Federal de Lavras, Lavras, Minas Gerais, Brazil.
+<sup>b</sup> Trichothecene genotype identified by the Luminex method: 15-ADON = 15-acetyl-deoxynivalenol; 3-ADON = 3-acetyl-deoxynivalenol; NIV = nivalenol.
+<sup>c</sup> Experiments: 1 = Perithecial production; 2 = Mycelial growth; 3 = Sporulation and germination; 4 = Pathogenicity; 5 = Sensitivity to tebuconazole.
+
 ### Perithecial production
@@
 The method was adapted from Chen & Zhou (2009).
-Briefly, 20g of grain was soaked in 10 mL distilled water and autoclaved for 20 min at 127 ° C daily during consecutive days.
-Afterwards, kernels were inoculated with three discs (6 mm diameter) of mycelia and kept in the dark for 21 days at 25 °C.
-Then, the colonized grains were transferred to a bench in a greenhouse with controlled temperature (25 ±3 °C) where they remained for seven days for drying.
+Briefly, 20g of grain was soaked in 10 mL distilled water and autoclaved for 20 min at 127 <sup>o</sup> C daily during consecutive days.
+Afterwards, kernels were inoculated with three discs (6 mm diameter) of mycelia and kept in the dark for 21 days at 25 <sup>o</sup>C.
+Then, the colonized grains were transferred to a bench in a greenhouse with controlled temperature (25 ±3 <sup>o</sup>C) where they remained for seven days for drying.
 After drying, they were placed in plastic box (11 x 11 x 3.5 cm) containing sterile sand layer (2 cm) and three sheets of filter paper on top of the sand.
@@
 
-Isolates grown on PDA (20 mL per dish) for seven days at 25 °C in continuous darkness provided mycelial discs (6 mm diameter) which were individually placed in the center of plastic Petri dishes (90 mm diameter).
-The cultures were incubated in growth chambers at two regimes of sub-optimal constant temperature (15 °C or 30 °C) under continuous darkness for five days when the diameter of the colony was measured from two perpendicular directions.
+Isolates grown on PDA (20 mL per dish) for seven days at 25 <sup>o</sup>C in continuous darkness provided mycelial discs (6 mm diameter) which were individually placed in the center of plastic Petri dishes (90 mm diameter).
+The cultures were incubated in growth chambers at two regimes of sub-optimal constant temperature (15 <sup>o</sup>C or 30 <sup>o</sup>C) under continuous darkness for five days when the diameter of the colony was measured from two perpendicular directions.
 The mycelial growth rate (mm/day) was calculated after subtracting the mycelium plug size (Spolti et al.
@@
 The asexual fecundity of the strains was evaluated based on the macroconidia production in culture medium.
-Isolates were grown on SNA media for seven days under a 12/12 h light/dark cycle at 25 ° C (Leslie and Summerell, 2006).
+Isolates were grown on SNA media for seven days under a 12/12 h light/dark cycle at 25 <sup>o</sup> C (Leslie and Summerell, 2006).
 After this period, three discs of 6 mm in diameter were removed from the edges of the developing colonies and immersed in 5 mL of sterile distilled water containing 0.1 mL Tween 0.001% in a test tube and shaken for 20 s.
@@
 2012b).
-After inoculation, the plants were kept at constant temperature (25 °C) to induce infection.
+After inoculation, the plants were kept at constant temperature (25 <sup>o</sup>C) to induce infection.
 The proportion of diseased spikelets in the wheat head was evaluated every three to five days during three weeks after inoculation.
@@
 One mycelia agar disc (6 mm in diameter) obtained from the edge of a 7-day-old culture was placed at the central position of a Petri dish (90 mm) containing 15 mL of PDA amended with the fungicide at each concentration.
-After four days of incubation at 25°C in darkness, radial growth was measured in two perpendicular directions and the disc diameter was subtracted.
+After four days of incubation at 25<sup>o</sup>C in darkness, radial growth was measured in two perpendicular directions and the disc diameter was subtracted.
 For each repeat of the experiment, two replicates were used.
@@
 ### Saprophytic fitness
+
 The proportion of grains with scores >1 (or >10% kernel area covered with perithecia) was apparently higher for maize and rice kernels substrates compared with wheat and for the 15-ADON isolates compared with NIV and 3-ADON (Figure 1A).
@@
 
-As expected, mean values of mycelial growth rate (MGR) was 84% higher at 30 °C (0.7 mm/day) than at 15 °C (1.2 mm/day) regardless of the trichothecene genotype.
+![](./figs/figure1.png)
+
+**Figure 1.** Relative frequency of perithecial production scores based on a visual ordinal (0-3) scale for the proportion of grains in each score for the three different cereal based substrates (wheat, rice or maize kernels) (A) and the least square means and 95% confidence interval for the perithecial production index which is a normalization of the frequency of the scores (B).
+
+As expected, mean values of mycelial growth rate (MGR) was 84% higher at 30 <sup>o</sup>C (0.7 mm/day) than at 15 <sup>o</sup>C (1.2 mm/day) regardless of the trichothecene genotype.
 Within each of the temperatures, there were no differences in MGR means among the trichothecene genotypes (*P* > 0.20) as well as between species within trichothecene genotypes (Figure 2A).
@@
 
-All strains yielded macroconidia and there were differences in conidia production among the genotypes (P = 0.057), which were more evident when species-genotypes were compared (P < 0.001).
+All strains yielded macroconidia and there were differences in conidia production among the genotypes (*P* = 0.057), which were more evident when species-genotypes were compared (*P* < 0.001).
 The most prolific species was *F.
@@
 austroamericanum for which the two strains differed by >70%.
-Therefore, the effect of genotype was not significant (P = 0.17), but when species-genotype was compared, the mean germination rate of F.
+Therefore, the effect of genotype was not significant (*P* = 0.17), but when species-genotype was compared, the mean germination rate of F.
 meridionale (10.5%) differed significantly from *F.
@@
 cortaderiae NIV (52.8%).
+
+![](./figs/figure2.png)
+
+**Figure 2.** Mycelial growth and spore production by strains of four Fusarium graminearum species complex members possessing a nivalenol (NIV) or an acetylate form of deoxynivalenol (DON) genotype (3-ADON or 15-ADON) assessed based on mycelial growth at 15<sup>o</sup> C or 30<sup>o</sup> C (A), macroconidia production (B) and conidia germination (C).
 
@@
 No formal test was used to compare these two cultivars because they were evaluated in different experiments.
+
 Results of the mixed model to test the effect of genotype in the mean area under the severity curve showed the 15-ADON *F.
@@
 
+![](./figs/figure3.png)
+
+**Figure 3.** Mean (standard deviation) severity of Fusarium head blight (FHB) in a moderately resistant cultivar, BRS Guamirim (A), and a susceptible cultivar, BRS 194 (B), inoculated with strains of different Fusarium graminearum species complex members possessing a nivalenol (NIV) or an acetylate form of deoxynivalenol (DON) genotype (3-ADON or 15-ADON) found associated with FHB in Brazil.
+
 ### Tebuconazole sensitivity
@@
 The estimated EC50 values ranged from 0.007 to 1.42 µg/mL across the strains.
-The highest mean EC50 value was found for one 15-ADON strain (1.42 µg/mL), while for the others the values were < 0.5 µg/mL.
+The highest mean EC50 value was found for one 15-ADON strain (1.42 µg/mL), while for the others the values were < 0.5 µg/mL (Fig.
+4).
 Differences among the genotypes were not apparent across the strains and the limited number of samples prevented formal comparison.
+
+![](./figs/figure4.png)
+
+Figure 4.
+The effective concentration of tebuconazole that reduces 50% (EC50) of the mycelial growth of four Fusarium graminearum species complex members possessing a nivalenol (NIV) or an acetylate form of deoxynivalenol (DON) trichothecene genotype (3-ADON or 15-ADON) found associated with Fusarium head blight of wheat in Brazil.
 
@@
 
-Table 1.
-Information for a collection of 30 strains belonging to four phylogenetic species of the *Fusarium graminearum* species complex isolated from symptomatic wheat heads from fields in Rio Grande do Sul State, Brazil, which were used in five different experiments in this study
+## Supplemental material
 
-Code | Location | 	Year  of isolation | 	Species |	Genotype |	Experiment
+![](./supp/figure1s.jpg)
 
-3065	| Panambi	2009	| *F. graminearum* | 	15-ADON	| 2, 3, 4 
-3066	| Ijuí	2009 |	*F. graminearum*	| 15-ADON	| 1,4
-3067	Carazinho	2010	*F. graminearum*	15-ADON	1,2,3,4
-3071	Tapejara	2010	*F. graminearum*	15-ADON	1,2,3,4,5
-3069	Coxilha	2010	*F. graminearum*	15-ADON	1, 2, 4,5
-3070	Santa Bárbara do Sul	2011	*F. graminearum*	15-ADON	1, 2, 4
-3830	Não-Me-Toque	2011	*F. graminearum*	15-ADON	1,2,3
-3833	Não-Me-Toque	2011	*F. graminearum*	15-ADON	1
-3064	Cruz Alta	2007	*F. graminearum*	15-ADON	1,4,5
-3068	Ernestina	2007	*F. graminearum*	15-ADON	1,2,3,4,5
-3382	Caseiros	2010	F. austroamericanum	3-ADON	1,2,3,5
-3384	Barreto	2010	F. austroamericanum	3-ADON	1,2,3,4,5
-3379	Carazinho	2010	F. cortaderiae	3-ADON	1,2,3,4,5
-3845	Lagoa Vermelha	2010	F. cortaderiae	3-ADON	1,2,3,4
-3841	Lagoa Vermelha	2010	F. cortaderiae	3-ADON	1,2,3,4,5
-3831	Panambi	2009	F. cortaderiae	NIV	1,2,3,4,5
-3836	Coronel Barros	2009	F. cortaderiae	NIV	1,2,4
-3837	Coronel Barros	2009	F. cortaderiae	NIV	1,2,3,4,5
-3840	Mato Castelhano	2010	F. cortaderiae	NIV	1,2,4
-3834	Barreto	2010	F. cortaderiae	NIV	1,2,4
-3378	Vacaria	2010	F. cortaderiae	NIV	1,2,3,4,5
-3835	Ernestina	2007	F. cortaderiae	NIV	1,2,3,4
-3375	Santa Bárbara do Sul	2009	F. meridionale	NIV	1,2,4
-3843	Ernestina	2009	F. meridionale	NIV	1,2,3,4
-3344	Coxilha	2009	F. meridionale	NIV	1,2,3,4,5
-3381	Caseiros	2010	F. meridionale	NIV	1,2,3,4,5
-3839	Marau	2011	F. meridionale	NIV	1,2,3,4,5
-3377	Sertão	2011	F. meridionale	NIV	1,2,3,4
-3838	Nonoai	2007	F. meridionale	NIV	1,2,3,4
-3374	Nonoai	2007	F. meridionale	NIV	1,2,4
-
-a Culture collection code: CML = Coleção Micológica de Lavras, Universidade Federal de Lavras, Lavras, Minas Gerais, Brazil.
-b Trichothecene genotype identified by the Luminex method: 15-ADON = 15-acetyl-deoxynivalenol; 3-ADON = 3-acetyl-deoxynivalenol; NIV = nivalenol.
-c Experiments: 1 = Perithecial production; 2 = Mycelial growth; 3 = Sporulation and germination; 4 = Pathogenicity; 5 = Sensitivity to tebuconazole.
-
-Figure 1.
-Relative frequency of perithecial production scores based on a visual ordinal (0-3) scale for the proportion of grains in each score for the three different cereal based substrates (wheat, rice or maize kernels) (A) and the least square means and 95% confidence interval for the perithecial production index which is a normalization of the frequency of the scores (B).
-
-Figure 2.
-Mycelial growth and spore production by strains of four Fusarium graminearum species complex members possessing a nivalenol (NIV) or an acetylate form of deoxynivalenol (DON) genotype (3-ADON or 15-ADON) assessed based on mycelial growth at 15° C or 30° C (A), macroconidia production (B) and conidia germination (C).
-
-Figure 3.
-Mean (standard deviation) severity of Fusarium head blight (FHB) in a moderately resistant cultivar, BRS Guamirim (A), and a susceptible cultivar, BRS 194 (B), inoculated with strains of different Fusarium graminearum species complex members possessing a nivalenol (NIV) or an acetylate form of deoxynivalenol (DON) genotype (3-ADON or 15-ADON) found associated with FHB in Brazil.
-
-Figure 5.
-The effective concentration of tebuconazole that reduces 50% (EC50) of the mycelial growth of four Fusarium graminearum species complex members possessing a nivalenol (NIV) or an acetylate form of deoxynivalenol (DON) trichothecene genotype (3-ADON or 15-ADON) found associated with Fusarium head blight of wheat in Brazil.
-
-Supplemental material
-
-Figure 1S
-
+**Figure 1S**.
 Diagrammatic key with four scores representing a percent range for the grain surface covered with perithecia, for rice, maize and wheat substrates.

### 2017-12-18 code.Rmd
Several changes including .doc as output

@@
-% Code
+% Analysis code
 
@@
 
-The data com from a series of different experiments.
-There is one single `xlsx` file with each experiment in a single sheet.
-Follows the commented code used for these analyses separated per experiment.
+The data were obtained from a series of randomized replicated experiments.
+There is one single `csv` file that contains data from each experiment.
+All analyses are presented below and separated for each experiment.
+I follow the steps (not in that exactly order) ilustrated in the R for Data Science book (http://r4ds.had.co.nz/) (Grolemund and Wickham 2017).
+These are: data import and tidy, explore (transform, visualize and model) and communicate.
 
-## Load packages
+These are the packages used.
 
@@
 
-Let's first import the data and have a look at the raw data.
+Let's have a look at the raw data.
 
@@
 
+The main response variable is ppi (perithecia production index), which was already calculated (it is an index based on the frequency of the scores).
+However, the original scores are included and consist of the number of grains on each score (0 to 3).
+
 ### Transform
 
-The data were prepared in the wide format where each score for percent perithecia coverage (0 -3) is in a separate column.
-Let's transform this dataframe from the wide to the long format using `gather` function of `dplyr`.
-We need to create a column for the variable name and another for the score.
+The frequecy of each score is shown in separate columns.
+To plot and further work with the data, we will need to format the data from the wide to the long format, where each frequency of the scores need to be in a single colum.
+We will use the `gather` function of `dplyr`.
+For this, we need to inform the name of columns for the score and the frequency.
 
@@
 
-Let's visualize the frequency of the scores for each species-genotype separated for each substrate. This is a nice plot to show in the paper!
+We now can visualize the frequency of the scores for each species-genotype separated (facet) for each substrate. This is a nice plot to have in the paper!
 
@@
 
-Since we have one species which has two genotypes, it may be important, besides testing the effect of species or genotype in the model, to create a species*genotype interaction variable and fit a mixed model to test the effect of this variable on the perithecial production index (ppi).
-This way the two genotypes within one species can be compared between them and with the others.
+Since we have one species with two genotypes, it may be important, besides testing the effect of species or genotype in the model, to create a species_genotype variable.
+We will then fit a mixed model to test the effect species_genotype on the perithecial production index (ppi).
+In this way the two genotypes within one species can be compared between them and with the others.
+We use the `unite` function.
 
@@
 
-Now we fit the model using `lmer` function. We treat `Isolate` as a random effects.
+Now we fit the mixed model using `lmer` function. We treat `Isolate` as a random effects.
 
 ```{r message=FALSE, warning=FALSE, paged.print=FALSE}
-glmm_per <- lmer(
+library(lme4)
+lmm_per <- lmer(
   ppi ~ species_genotype * substrate + (1 | Isolate),
@@
 
-Let's check the single and interaction effects.
+Let's check the single and interaction effects using `Anova` function of the `car` package.
 
 ```{r message=FALSE, warning=FALSE}
-Anova(glmm_per)
+library(car)
+Anova(lmm_per)
 ```
@@
 ```{r}
-plot(glmm_per, type = c("p", "smooth"))
-qqmath(glmm_per, id = 0.05)
+plot(lmm_per, type = c("p", "smooth"))
+qqmath(lmm_per, id = 0.05)
 ```
@@
 Let's now compare the means of treatments and create a dataframe which will be use to further create a plot with the estimated means and confidence interval.
-We used the `emmeans` package which is an update for the old `lsmeans` package.
+We will use the `emmeans` package which is an update for the old `lsmeans` package.
 The syntax is the same as before.
@@
 ```{r message=FALSE, warning=FALSE}
-medias <- emmeans(glmm_per, ~ species_genotype * substrate)
+library(emmeans)
+medias <- emmeans(lmm_per, ~ species_genotype * substrate)
 med <- cld(medias, Letters = LETTERS, alpha = .05)
@@
 
-This new `med` dataframe does not contain the two original variables we want to use to plot the data. Let's separate the `species_genotype` variable and create `genotype` and `species`.
+This new `med` dataframe does not contain the two original variables we want to use to plot the data.
+Let's separate the `species_genotype` variable and create `genotype` and `species`.
+We use the `separate` function.
 
@@
 
-We now produce the plot with point estimate for the means and error bars for the 95%CI of the `ppi` for each `genotype`. We add `species` to the legend when using `color` argument.
+We now can plot the point estimates for the means and respective 95%CI of the `ppi` for each `genotype`. We add `species` legends when using `color` argument so we can have five species_genotype.
 
@@
 
-There is one `csv` file with data from a single experiment. We will use `read_csv` to impor the data.
+The main variable is mycelia growth rate, or the ratio of the growth at the fifth day and number of days. Besides the species and genotype factors, there are two levels of temperatures being tested.
 
@@
 
-We will now produce the plot with the means of mycelial growth rate for each species-genotype and temperature. This will be the final plot for publication.
+We can plot means of mycelial growth rate for each species-genotype and temperature as facetting variable. This plot will be included in the publication.
 
@@
 
-We now create different datasets for each temperature and test the effect of `species_genotype` within each temperature.
+The interaction was significant. We now create different datasets for each temperature and test the effect of `species_genotype` within each temperature.
 
@@
 
-## Sporoulation & Germination
+## Sporulation & Germination
 
-In this experiment, the production of macroconidia and further germination of 20 randomly selected spores of the different strains of the species-genotype combination, were evaluated in two runs of the experiment.
+In this experiment, the production of macroconidia and further germination of 20 randomly selected spores of the different strains of the species-genotype combination, was evaluated in two runs of the experiment.
 
@@
 
-Anova model
+Mixed model for testing the effect of genotype.
 
 ```{r message=FALSE, warning=FALSE}
-lmer_spor <- lmer(spores_ml ~ species_genotype + (1 | species_genotype), data = spor, REML = FALSE)
+lmer_spor <- lmer(spores_ml ~ genotype + (1 | genotype / isolate), data = spor, REML = FALSE)
+
+```
+
+```{r}
+plot(lmer_spor, type = c("p", "smooth"), col = "black")
+qqmath(lmer_spor, id = 0.05, col = "black")
+```
+
+Model fitted well the data. Let's proceed.
+
+```{r}
 Anova(lmer_spor)
-plot(lmer_spor, type = c("p", "smooth"))
-qqmath(lmer_spor, id = 0.05)
 ```
@@
 ```{r}
-medias <- emmeans(lmer_spor, ~species_genotype)
+medias <- emmeans(lmer_spor, ~genotype)
 med <- cld(medias, Letters = LETTERS, alpha = .05)
@@
 ### Figure 2
+
+We will produce a combo figure for the three variables.
 
@@
 
-## Pathogenicity tests
+## Pathogenicity
+
+The pathogenicity was assessed based on the progress of the symptoms from the central (inoculated) spikelet.
+The data was transformed to percent severity of the spike (proportion of diseased spikelets).
+There is one obsevation of severity for each day after inoculation (dai) and individual spike.
 
@@
 pathogen <- read_csv(here("data", "pathogenicity.csv")) %>%
-  filter(isolate != "test")
+  filter(isolate != "test") %>% 
+  select (-c(8:12))
+
+pathogen
 ```
@@
 
-Let's prepare several summary statistics for the disease variables per cultivar. We will calculate the mean and standard deviation of proportion of diseased spikelets.
+Since we have spikes as replicates, we will calculate the mean and standard deviation of severity for plotting purposes.
 
@@
   filter(cultivar == "BRS 194") %>%
-  select(1:7, 9, 11, 12, 13) %>%
-  gather(position, count, 8:9) %>%
-  # filter(exp == 2) %>%
   group_by(cultivar, dai, genotype, species) %>%
@@
   filter(cultivar == "BRS Guamirim") %>%
-  select(1:7, 9, 11:13) %>%
-  gather(position, count, 8:9) %>%
    # filter(exp == 2) %>%
-  group_by(cultivar, dai, genotype, species, position) %>%
+  group_by(cultivar, dai, genotype, species) %>%
   summarize(
@@
 To make our life easier, lets use the `do` in conjuction with the `tidy` functions of the `dplyr` and `broom` package, respectively.
-This will calculate the audpce for each spike.
+This will calculate the audpc for each spike.
+
+#### BRS 194
 
@@
 
+Now we can plot the means and standard deviation for each isolate.
+
+```{r}
+audpc_194 %>%
+  group_by(isolate, species, genotype) %>%
+  summarize(
+    mean_audpc = mean(audpc),
+    sd_audpc = sd(audpc)
+  ) %>%
+  ggplot(aes(reorder(isolate, mean_audpc), mean_audpc, color = genotype, shape = species)) +
+  coord_flip() +
+  theme_few() +
+  geom_point(position = position_dodge(width = 0.4)) +
+  geom_errorbar(aes(min = mean_audpc - sd_audpc, max = mean_audpc + sd_audpc), position = position_dodge(width = 0.4), width = 0.2)
+```
+
+#### BRS Guamirim
+
 ```{r}
@@
 
+Now we can plot the means and standard deviation for each isolate.
+
 ```{r}
-audpc_194 %>%
-  group_by(species, genotype) %>%
+audpc_gua %>%
+  group_by(isolate, species, genotype) %>%
   summarize(
@@
   ) %>%
-  ggplot(aes(species, mean_audpc, color = genotype)) +
+  ggplot(aes(reorder(isolate, mean_audpc), mean_audpc, color = genotype, shape = species)) +
   coord_flip() +
@@
 
-```{r}
-audpc_gua %>%
-  group_by(species, genotype) %>%
-  summarize(
-    mean_audpc = mean(audpc),
-    sd_audpc = sd(audpc)
-  ) %>%
-  ggplot(aes(species, mean_audpc, color = genotype)) +
-  coord_flip() +
-  theme_few() +
-  geom_point(position = position_dodge(width = 0.4)) +
-  geom_errorbar(aes(min = mean_audpc - sd_audpc, max = mean_audpc + sd_audpc), position = position_dodge(width = 0.4), width = 0.2)
-```
+### Mixed model
 
-### Model
+We fit a separate model for each cultivar because the experiments were conducted at different times.
+We fit a multilevel to test the effect of specis_genotype as fixed effects and spikes nested within isolates as random effects.
+The species_genotype interaction was non-significant (not shown), and then a simpler model is used.
 
@@
 ```{r}
+mix_gua <- lmer(
+  audpc ~ species_genotype  +
+    (1 | isolate/spike),
+  data = audpc_gua, REML = FALSE)
 
-mix_gua <- lmer(
-  audpc ~ genotype +
-    (1 | exp / spike),
-  data = audpc_gua, REML = FALSE
-)
+mix_gua
+
 ```
@@
 ```{r message=FALSE, warning=FALSE}
-
 plot(mix_gua, type = c("p", "smooth"), col = "black")
@@
 
+The model seems OK. Let's check the significance of the species_genotype and estimated means.
+
 ```{r}
 Anova(mix_gua)
-
-means <- emmeans(mix_gua, ~ genotype)
+means <- emmeans(mix_gua, ~ species_genotype)
 cld(means)
@@
 
+We can see that the variation among isolates was too large and so no difference could be detected among the five species_genotype in the moderate resistant cultivar.
+
+#### BRS 194
+
+We now apply the same model for the susceptible cultivar.
+
 ```{r}
-mix_194 <- lmer(log(audpc) ~ genotype + (1 | exp / spike), data = audpc_194, REML = FALSE)
+mix_194 <- lmer(audpc ~ species_genotype + 
+                  (1 | isolate/spike), 
+                data = audpc_194)
+mix_194
 ```
 
-Evaluate the model
+Evaluate the model.
 
 ```{r}
-
 plot(mix_194, type = c("p", "smooth"), col = "black")
@@
 
-#### BRS 194
+Note: I used both the original and the log-transformed `audpc` and the results were the same. Hence, I kept the original data.
 
@@
 Anova(mix_194)
-means <- emmeans(mix_194, ~ genotype)
+means <- emmeans(mix_194, ~ species_genotype)
 cld(means)
@@
 
+Again, the effect of species_genotype was not significant due to large variation among isolates.
+
 ## Fungicide sensitivity
+
+The response variable is the EC50 estimated in two experiments (true replicates) for each selected isolate from each species_genotype.
 
@@
 ec50 <- read_csv(here("data", "ec50.csv"))
+ec50
 ```
@@
 
-```{r}
+Let's obtain the mean and standard deviation of ec50 for each strain.
+
+```{r message=FALSE, warning=FALSE}
 ec502 <- ec50 %>%
-  group_by(isolate, species, genotype) %>%
+   unite(species_genotype, species, genotype, sep = "_", remove = F) %>% 
+  group_by(isolate, species, genotype, species_genotype) %>%
   summarize(
@@
   )
+ec502
 ```
@@
 
-Let's have a look at the mean EC50 for the isolates
+Let's have a look at the mean EC50 for each isolate.
 
-```{r}
+```{r message=FALSE, warning=FALSE}
 ec502 %>%
-  ggplot(aes(genotype, mean_ec50, color = species)) +
-  geom_jitter(width = 0.1) +
-  # geom_errorbar(aes(min = mean_ec50 - sd_ec50, max = mean_ec50 + sd_ec50), width=0.2, position = position_dodge(width = 0.2))+
+  ggplot(aes(reorder(isolate, mean_ec50), mean_ec50, shape = species, color = genotype)) +
+  geom_point() +
+  geom_errorbar(aes(min = mean_ec50 - sd_ec50, max = mean_ec50 + sd_ec50), width = 0.2)+
   theme_few() +
+  coord_flip()+
   theme(legend.position = "right", legend.text = element_text(size = 7, face = "italic")) +
@@
   scale_color_viridis(discrete = TRUE) +
-  labs(y = (expression(paste("EC"[50], " (", mu, "g/ml)"))), x = "Trichothecene genotype", color = "Species") +
+  labs(y = (expression(paste("EC"[50], " (", mu, "g/ml)"))), x = "FGSC strain", shape = "Species", color = "Genotype" ) +
   ggsave(here("figs", "ec50-new.png"), width = 4.5, height = 3)

### 2017-12-18 index.Rmd
Several changes including .doc as output

@@
 
-<img src = "https://www.apsnet.org/edcenter/intropp/lessons/fungi/ascomycetes/PublishingImages/Fusariumdiscyclesm.jpg" align= right width = 250> This is a research compendium that contains and describe the data and the R codes for analyzing and visualizing these data obtained from a series of experiments conducted as an MSc research in the Del Ponte lab.
-The main research's objective was to assess and compare members of a species complex (FGSC), with varying toxigenic potential, that cause [Fusarium head blight or scab](https://www.apsnet.org/edcenter/intropp/lessons/fungi/ascomycetes/Pages/Fusarium.aspx), an important disease of wheat and barley worldwide.
+<img src = "https://www.apsnet.org/edcenter/intropp/lessons/fungi/ascomycetes/PublishingImages/Fusariumdiscyclesm.jpg" align= right width = 270> This research compendium  contains the data and the commented R codes for analyzing and visualizing these data obtained from a series of experiments conducted as an MSc research in the Del Ponte lab.
+
+The main objective was to assess and compare members of a species complex (FGSC), with varying toxigenic potential, which cause [Fusarium head blight (FHB) or scab](https://www.apsnet.org/edcenter/intropp/lessons/fungi/ascomycetes/Pages/Fusarium.aspx).
+FHB is an important disease of wheat and barley worldwide.
 Several fitness-related traits were compared in five experiments conducted both *in vitro* and in *planta*.

### 2017-12-18 manuscript.Rmd
Several changes including .doc as output

@@
 title: "Fitness traits of deoxynivalenol and nivalenol-producing Fusarium graminearum species complex strains from wheat"
+author: "C. P. Nicolli, F. J. Machado, P. Spolti, E. M. Del Ponte"
+csl: plant-disease.csl
 bibliography: crossref.bib
-csl: plant-disease.csl
+output: 
+  word_document:
+    reference_docx: manuscript.docx
 ---
 
-Camila P. Nicolli, Franklin J. Machado, Piérri Spolti, Emerson M. Del Ponte   
 Departamento de Fitopatologia, Universidade Federal de Viçosa, 36570-000, Viçosa, MG, Brazil.
@@
 However, 3-ADON and nivalenol(NIV) chemotypes have been found in other members of the species complex causing FHB in wheat.
-To improve our understanding of the pathogen ecology, we assessed a range of fitness-related traits in a sample of 30 strains representatives of 15-ADON (**F.
-graminearum**), 3-ADON (*F.
+To improve our understanding of the pathogen ecology, we assessed a range of fitness-related traits in a sample of 30 strains representatives of 15-ADON (*F.
+graminearum*), 3-ADON (*F.
 cortaderiae* and *F.
@@
 
-Key words: Triticum aestivum, trichothecenes, fitness, FGSC.
+Key words: *Triticum aestivum*, trichothecenes, fitness, FGSC.
 
@@
 2015).
-The **F.
-graminearum** strains found in Brazilian wheat and barley analyzed so far are exclusively of the 15-acetyl(A)DON genotype, a mainly DON producer (Astolfi et al.
+The *F.
+graminearum* strains found in Brazilian wheat and barley analyzed so far are exclusively of the 15-acetyl(A)DON genotype, a mainly DON producer (Astolfi et al.
 2012; Castañares et al.
@@
 Empirical knowledge of some components of the life cycle of other FGSC members than *F.
-graminearum* and F.
-meridionale in Brazil, such as the saprophytic ability, and whether this is influenced by the cereal-based substrate and temperature, which may provide competitive advantages for a species to build up inoculum for further spread are not available.
+graminearum* and *F.
+meridionale* in Brazil, such as the saprophytic ability, and whether this is influenced by the cereal-based substrate and temperature, which may provide competitive advantages for a species to build up inoculum for further spread are not available.
 In this study, a sample of 30 strains representative of three trichothecene genotypes among four FGSC members associated with FHB of wheat in southern Brazil was selected to investigate the fitness-related traits related to both the saprophytic and pathogenic phase.
@@
 A total of five different experiments were conducted to assess saprophytic (sexual and asexual reproduction), pathogenic (severity towards two cultivars), and sensitivity to a triazole fungicide (tebuconazole).
-
-**Table 1.** Information for a collection of 30 strains belonging to four phylogenetic species of the Fusarium graminearum species complex isolated from symptomatic wheat heads from fields in Rio Grande do Sul State, Brazil, which were used in five different experiments in this study
-
-| CML Code<sup>a</sup> | Location             | Year of isolation | Species             | Genotype<sup>b</sup> | Experiment<sup>c</sup> |
-|------|----------------------|-------------------|---------------------|----------|------------|
-| 3065 | Panambi              | 2009              | F. graminearum      | 15-ADON  | 2, 3, 4    |
-| 3066 | Ijuí                 | 2009              | F. graminearum      | 15-ADON  | 1,4        |
-| 3067 | Carazinho            | 2010              | F. graminearum      | 15-ADON  | 1,2,3,4    |
-| 3071 | Tapejara             | 2010              | F. graminearum      | 15-ADON  | 1,2,3,4,5  |
-| 3069 | Coxilha              | 2010              | F. graminearum      | 15-ADON  | 1, 2, 4,5  |
-| 3070 | Santa Bárbara do Sul | 2011              | F. graminearum      | 15-ADON  | 1, 2, 4    |
-| 3830 | Não-Me-Toque         | 2011              | F. graminearum      | 15-ADON  | 1,2,3      |
-| 3833 | Não-Me-Toque         | 2011              | F. graminearum      | 15-ADON  | 1          |
-| 3064 | Cruz Alta            | 2007              | F. graminearum      | 15-ADON  | 1,4,5      |
-| 3068 | Ernestina            | 2007              | F. graminearum      | 15-ADON  | 1,2,3,4,5  |
-| 3382 | Caseiros             | 2010              | F. austroamericanum | 3-ADON   | 1,2,3,5    |
-| 3384 | Barreto              | 2010              | F. austroamericanum | 3-ADON   | 1,2,3,4,5  |
-| 3379 | Carazinho            | 2010              | F. cortaderiae      | 3-ADON   | 1,2,3,4,5  |
-| 3845 | Lagoa Vermelha       | 2010              | F. cortaderiae      | 3-ADON   | 1,2,3,4    |
-| 3841 | Lagoa Vermelha       | 2010              | F. cortaderiae      | 3-ADON   | 1,2,3,4,5  |
-| 3831 | Panambi              | 2009              | F. cortaderiae      | NIV      | 1,2,3,4,5  |
-| 3836 | Coronel Barros       | 2009              | F. cortaderiae      | NIV      | 1,2,4      |
-| 3837 | Coronel Barros       | 2009              | F. cortaderiae      | NIV      | 1,2,3,4,5  |
-| 3840 | Mato Castelhano      | 2010              | F. cortaderiae      | NIV      | 1,2,4      |
-| 3834 | Barreto              | 2010              | F. cortaderiae      | NIV      | 1,2,4      |
-| 3378 | Vacaria              | 2010              | F. cortaderiae      | NIV      | 1,2,3,4,5  |
-| 3835 | Ernestina            | 2007              | F. cortaderiae      | NIV      | 1,2,3,4    |
-| 3375 | Santa Bárbara do Sul | 2009              | F. meridionale      | NIV      | 1,2,4      |
-| 3843 | Ernestina            | 2009              | F. meridionale      | NIV      | 1,2,3,4    |
-| 3344 | Coxilha              | 2009              | F. meridionale      | NIV      | 1,2,3,4,5  |
-| 3381 | Caseiros             | 2010              | F. meridionale      | NIV      | 1,2,3,4,5  |
-| 3839 | Marau                | 2011              | F. meridionale      | NIV      | 1,2,3,4,5  |
-| 3377 | Sertão               | 2011              | F. meridionale      | NIV      | 1,2,3,4    |
-| 3838 | Nonoai               | 2007              | F. meridionale      | NIV      | 1,2,3,4    |
-| 3374 | Nonoai               | 2007              | F. meridionale      | NIV      | 1,2,4      |
-
-<sup>a</sup> Culture collection code: CML = Coleção Micológica de Lavras, Universidade Federal de Lavras, Lavras, Minas Gerais, Brazil.
-<sup>b</sup> Trichothecene genotype identified by the Luminex method: 15-ADON = 15-acetyl-deoxynivalenol; 3-ADON = 3-acetyl-deoxynivalenol; NIV = nivalenol.
-<sup>c</sup> Experiments: 1 = Perithecial production; 2 = Mycelial growth; 3 = Sporulation and germination; 4 = Pathogenicity; 5 = Sensitivity to tebuconazole.
 
@@
 
-![](./figs/figure1.png)
-
-**Figure 1.** Relative frequency of perithecial production scores based on a visual ordinal (0-3) scale for the proportion of grains in each score for the three different cereal based substrates (wheat, rice or maize kernels) (A) and the least square means and 95% confidence interval for the perithecial production index which is a normalization of the frequency of the scores (B).
-
 As expected, mean values of mycelial growth rate (MGR) was 84% higher at 30 <sup>o</sup>C (0.7 mm/day) than at 15 <sup>o</sup>C (1.2 mm/day) regardless of the trichothecene genotype.
@@
 
-![](./figs/figure2.png)
-
-**Figure 2.** Mycelial growth and spore production by strains of four Fusarium graminearum species complex members possessing a nivalenol (NIV) or an acetylate form of deoxynivalenol (DON) genotype (3-ADON or 15-ADON) assessed based on mycelial growth at 15<sup>o</sup> C or 30<sup>o</sup> C (A), macroconidia production (B) and conidia germination (C).
-
 ### Pathogenicity
@@
 
-![](./figs/figure3.png)
-
-**Figure 3.** Mean (standard deviation) severity of Fusarium head blight (FHB) in a moderately resistant cultivar, BRS Guamirim (A), and a susceptible cultivar, BRS 194 (B), inoculated with strains of different Fusarium graminearum species complex members possessing a nivalenol (NIV) or an acetylate form of deoxynivalenol (DON) genotype (3-ADON or 15-ADON) found associated with FHB in Brazil.
-
 ### Tebuconazole sensitivity
@@
 Differences among the genotypes were not apparent across the strains and the limited number of samples prevented formal comparison.
-
-![](./figs/figure4.png)
-
-Figure 4.
-The effective concentration of tebuconazole that reduces 50% (EC50) of the mycelial growth of four Fusarium graminearum species complex members possessing a nivalenol (NIV) or an acetylate form of deoxynivalenol (DON) trichothecene genotype (3-ADON or 15-ADON) found associated with Fusarium head blight of wheat in Brazil.
 
@@
 
-Spolti, P., Del PontE, E.M.
+Spolti, P., Del Ponte, E.M.
 2013.
@@
 2014b.
-Triazole sensitivity in a contemporary population of Fusarium graminearum from New York wheat and competitiveness of a tebuconazole-resistant Isolate.
+Triazole sensitivity in a contemporary population of *Fusarium graminearum* from New York wheat and competitiveness of a tebuconazole-resistant Isolate.
 Plant Dis.
@@
 
-Tunali, B., Obanor, F., Erginbas, G., Westecott, R.A., Nicol, J., Chakraborty, S. 2012. Fitness of three Fusarium pathogens of wheat. Microbiol Ecol. 81:596-609.
+Tunali, B., Obanor, F., Erginbas, G., Westecott, R.A., Nicol, J., Chakraborty, S. 2012. Fitness of three *Fusarium* pathogens of wheat. Microbiol Ecol. 81:596-609.
 
@@
 
+**Table 1.** Information for a collection of 30 strains belonging to four phylogenetic species of the *Fusarium graminearum* species complex isolated from symptomatic wheat heads from fields in Rio Grande do Sul State, Brazil, which were used in five different experiments in this study
+
+| CML <sup>a</sup> | Location             | Year  | Species             | Genotype<sup>b</sup> | Experiment<sup>c</sup> |
+|------|----------------------|-------------------|---------------------|----------|------------|
+| 3065 | Panambi              | 2009              | F. graminearum      | 15-ADON  | 2, 3, 4    |
+| 3066 | Ijuí                 | 2009              | F. graminearum      | 15-ADON  | 1,4        |
+| 3067 | Carazinho            | 2010              | F. graminearum      | 15-ADON  | 1,2,3,4    |
+| 3071 | Tapejara             | 2010              | F. graminearum      | 15-ADON  | 1,2,3,4,5  |
+| 3069 | Coxilha              | 2010              | F. graminearum      | 15-ADON  | 1, 2, 4,5  |
+| 3070 | Santa Bárbara do Sul | 2011              | F. graminearum      | 15-ADON  | 1, 2, 4    |
+| 3830 | Não-Me-Toque         | 2011              | F. graminearum      | 15-ADON  | 1,2,3      |
+| 3833 | Não-Me-Toque         | 2011              | F. graminearum      | 15-ADON  | 1          |
+| 3064 | Cruz Alta            | 2007              | F. graminearum      | 15-ADON  | 1,4,5      |
+| 3068 | Ernestina            | 2007              | F. graminearum      | 15-ADON  | 1,2,3,4,5  |
+| 3382 | Caseiros             | 2010              | F. austroamericanum | 3-ADON   | 1,2,3,5    |
+| 3384 | Barreto              | 2010              | F. austroamericanum | 3-ADON   | 1,2,3,4,5  |
+| 3379 | Carazinho            | 2010              | F. cortaderiae      | 3-ADON   | 1,2,3,4,5  |
+| 3845 | Lagoa Vermelha       | 2010              | F. cortaderiae      | 3-ADON   | 1,2,3,4    |
+| 3841 | Lagoa Vermelha       | 2010              | F. cortaderiae      | 3-ADON   | 1,2,3,4,5  |
+| 3831 | Panambi              | 2009              | F. cortaderiae      | NIV      | 1,2,3,4,5  |
+| 3836 | Coronel Barros       | 2009              | F. cortaderiae      | NIV      | 1,2,4      |
+| 3837 | Coronel Barros       | 2009              | F. cortaderiae      | NIV      | 1,2,3,4,5  |
+| 3840 | Mato Castelhano      | 2010              | F. cortaderiae      | NIV      | 1,2,4      |
+| 3834 | Barreto              | 2010              | F. cortaderiae      | NIV      | 1,2,4      |
+| 3378 | Vacaria              | 2010              | F. cortaderiae      | NIV      | 1,2,3,4,5  |
+| 3835 | Ernestina            | 2007              | F. cortaderiae      | NIV      | 1,2,3,4    |
+| 3375 | Santa Bárbara do Sul | 2009              | F. meridionale      | NIV      | 1,2,4      |
+| 3843 | Ernestina            | 2009              | F. meridionale      | NIV      | 1,2,3,4    |
+| 3344 | Coxilha              | 2009              | F. meridionale      | NIV      | 1,2,3,4,5  |
+| 3381 | Caseiros             | 2010              | F. meridionale      | NIV      | 1,2,3,4,5  |
+| 3839 | Marau                | 2011              | F. meridionale      | NIV      | 1,2,3,4,5  |
+| 3377 | Sertão               | 2011              | F. meridionale      | NIV      | 1,2,3,4    |
+| 3838 | Nonoai               | 2007              | F. meridionale      | NIV      | 1,2,3,4    |
+| 3374 | Nonoai               | 2007              | F. meridionale      | NIV      | 1,2,4      |
+
+<sup>a</sup> Culture collection code: CML = Coleção Micológica de Lavras, Universidade Federal de Lavras, Lavras, Minas Gerais, Brazil.
+<sup>b</sup> Trichothecene genotype identified by the Luminex method: 15-ADON = 15-acetyl-deoxynivalenol; 3-ADON = 3-acetyl-deoxynivalenol; NIV = nivalenol.
+<sup>c</sup> Experiments: 1 = Perithecial production; 2 = Mycelial growth; 3 = Sporulation and germination; 4 = Pathogenicity; 5 = Sensitivity to tebuconazole.
+
+![](./figs/figure1.png)
+
+**Figure 1.** Relative frequency of perithecial production scores based on a visual ordinal (0-3) scale for the proportion of grains in each score for the three different cereal based substrates (wheat, rice or maize kernels) (A) and the least square means and 95% confidence interval for the perithecial production index which is a normalization of the frequency of the scores (B).
+
+![](./figs/figure2.png)
+
+**Figure 2.** Mycelial growth and spore production by strains of four Fusarium graminearum species complex members possessing a nivalenol (NIV) or an acetylate form of deoxynivalenol (DON) genotype (3-ADON or 15-ADON) assessed based on mycelial growth at 15<sup>o</sup> C or 30<sup>o</sup> C (A), macroconidia production (B) and conidia germination (C).
+
+![](./figs/figure3.png)
+
+**Figure 3.** Mean (standard deviation) severity of Fusarium head blight (FHB) in a moderately resistant cultivar, BRS Guamirim (A), and a susceptible cultivar, BRS 194 (B), inoculated with strains of different Fusarium graminearum species complex members possessing a nivalenol (NIV) or an acetylate form of deoxynivalenol (DON) genotype (3-ADON or 15-ADON) found associated with FHB in Brazil.
+
+![](./figs/figure4.png)
+
+Figure 4.
+The effective concentration of tebuconazole that reduces 50% (EC50) of the mycelial growth of four Fusarium graminearum species complex members possessing a nivalenol (NIV) or an acetylate form of deoxynivalenol (DON) trichothecene genotype (3-ADON or 15-ADON) found associated with Fusarium head blight of wheat in Brazil.
+
 ## Supplemental material

### 2017-12-20 data.Rmd
new style
Several changes including .doc as output
Added OSF link
Small changes

@@
 
-The data were obtained from five different (mostly repeated in time) experiments, and were organized in a single `xlsx` file.
+The data for this study were obtained from five different (mostly repeated in time) experiments, and were organized in a single `xlsx` file.
 Data from each experiment are in a separate sheet and were exported to a `csv` file.
-
-The data are already organized in a rectangular (tidy) format, mostly in the long-format (a raw is an observation), but some are in the wide (observations in multiple colums), and so will need some wrangling.
+They are already organized in a rectangular (tidy) format, mostly in the long-format (a raw is an observation).
+For some experiments, they are in the wide format (observations in multiple colums), and so will need some wrangling.
 This munging is done within the respective section of the data analysis page, after data import.
@@
 
-The data are stored in the `data` folder of the GitHub repository.
+> The data are stored in the following repositories:
 
-https://github.com/phytopathology/paper-FGSC-fitness/data
+https://github.com/phytopathology/paper-FGSC-fitness/tree/master/dat  
+https://osf.io/c2mbr/ 
+Identifier: DOI 10.17605/OSF.IO/C2MBR

### 2017-12-20 index.Rmd
Small changes

@@
 
-<img src = "https://www.apsnet.org/edcenter/intropp/lessons/fungi/ascomycetes/PublishingImages/Fusariumdiscyclesm.jpg" align= right width = 270> This research compendium  contains the data and the commented R codes for analyzing and visualizing these data obtained from a series of experiments conducted as an MSc research in the Del Ponte lab.
-
+<img src = "https://www.apsnet.org/edcenter/intropp/lessons/fungi/ascomycetes/PublishingImages/Fusariumdiscyclesm.jpg" align= right width = 270> This research compendium  contains the data and R codes for analyzing and visualizing these data obtained from a series of experiments conducted as an MSc research in the Del Ponte lab.
 The main objective was to assess and compare members of a species complex (FGSC), with varying toxigenic potential, which cause [Fusarium head blight (FHB) or scab](https://www.apsnet.org/edcenter/intropp/lessons/fungi/ascomycetes/Pages/Fusarium.aspx).
-FHB is an important disease of wheat and barley worldwide.
 Several fitness-related traits were compared in five experiments conducted both *in vitro* and in *planta*.
 Check the manuscript page.
+
+## Citation
+
+> Del Ponte EM. Fitness-related traits of *Fusarium graminearum* species complex: a research compendia. Online at https://phytopathology.github.io/paper-FGSC-fitness/
+
+## Pre-print version
+
+> BIORXIV/2017/235895
+Fitness traits of deoxynivalenol and nivalenol-producing *Fusarium graminearum* species complex strains from wheat
+Camila P Nicolli, Franklin J Machado, Pierri Spolti, and Emerson M Del Ponte  
+https://doi.org/10.1101/235895
 
@@
 
-## Citation
-
-Please cite this research compendium as:
-
-> Del Ponte EM. Fitness-related traits of Fusarium graminearum species complex: a research compendia. Online at https://phytopathology.github.io/paper-FGSC-fitness/
-
 ## Licences

### 2018-01-21 code.Rmd
Small changes
Check spell and text edit

@@
 All analyses are presented below and separated for each experiment.
-I follow the steps (not in that exactly order) ilustrated in the R for Data Science book (http://r4ds.had.co.nz/) (Grolemund and Wickham 2017).
-These are: data import and tidy, explore (transform, visualize and model) and communicate.
+I follow the steps (not in that exactly order) illustrated in the R for Data Science book (http://r4ds.had.co.nz/) (Grolemund and Wickham 2017).
+They are: data import and tidy, explore (transform, visualize and model) and communicate.
 
@@
 
-The main response variable is ppi (perithecia production index), which was already calculated (it is an index based on the frequency of the scores).
+The main response variable is ppi (perithecia production index), which was already calculated.
+It is an index based on the frequency of the scores given by given by: PPI = ((n0× 0)+(n1 × 1)+(n2 × 2)+(n3 × 3)) × 100) / (ntotal × 3) where nx = number of grains in the respective x category (x = 1 to 4 score) and ntotal = total number of grains (25 in this study).
 However, the original scores are included and consist of the number of grains on each score (0 to 3).
@@
 
-The frequecy of each score is shown in separate columns.
-To plot and further work with the data, we will need to format the data from the wide to the long format, where each frequency of the scores need to be in a single colum.
+The frequency of each score is shown in separate columns.
+To plot and further work with the data, we will need to reshape the data from the wide to the long format, where each frequency of the scores need to be in a single column.
 We will use the `gather` function of `dplyr`.
-For this, we need to inform the name of columns for the score and the frequency.
+For this, we need to inform the name of columns for the `score` and `frequency` variables.
 
@@
 
-We now can visualize the frequency of the scores for each species-genotype separated (facet) for each substrate. This is a nice plot to have in the paper!
+We can now visualize the frequency of the scores for each species-genotype separated (facet) for each substrate. This is a nice plot to have in the paper!
 
@@
 
-Since we have one species with two genotypes, it may be important, besides testing the effect of species or genotype in the model, to create a species_genotype variable.
+Since we have one species with two genotypes, it may be important, besides testing the effect of species or genotype in the model, to create a `species_genotype` variable.
 We will then fit a mixed model to test the effect species_genotype on the perithecial production index (ppi).
-In this way the two genotypes within one species can be compared between them and with the others.
-We use the `unite` function.
+In this way the two genotypes within one species can be compared between them and among the others.
+We use the `unite` function to create this variable.
 
@@
 
-Now we fit the mixed model using `lmer` function. We treat `Isolate` as a random effects.
+Now we can fit the mixed model using `lmer` function. We treat `Isolate` as a random effects.
 
@@
 ```{r}
-plot(lmm_per, type = c("p", "smooth"))
-qqmath(lmm_per, id = 0.05)
+plot(lmm_per, type = c("p", "smooth"), col = "black")
+qqmath(lmm_per, id = 0.05, col = "black")
 ```
@@
 Not too bad.
-Let's now compare the means of treatments and create a dataframe which will be use to further create a plot with the estimated means and confidence interval.
+Let's now compare the means of treatments and create a data-frame which will be use to further create a plot with the estimated means and confidence interval.
 We will use the `emmeans` package which is an update for the old `lsmeans` package.
@@
 
-This new `med` dataframe does not contain the two original variables we want to use to plot the data.
+This new `med` data-frame does not contain the two original variables we want to use to plot the data.
 Let's separate the `species_genotype` variable and create `genotype` and `species`.
@@
 grid1 <- plot_grid(g1, g2, labels = c("A", "B"), align = "hv", ncol = 1)
-ggsave(here("figs", "figure1.png"), grid1, width = 6, height = 7)
+ggsave(here("figs", "figure1.png"), grid1, dpi = 600, width = 6, height = 7)
 grid1
@@
 
-We can plot means of mycelial growth rate for each species-genotype and temperature as facetting variable. This plot will be included in the publication.
+We can plot means of mycelial growth rate for each species-genotype and temperature as faceting variable. This plot will be included in the publication.
 
@@
   theme(
-    legend.position = "none",
+    legend.position = "top",
     legend.text = element_text(size = 7, face = "italic")
@@
 
-The interaction was significant. We now create different datasets for each temperature and test the effect of `species_genotype` within each temperature.
+The interaction was significant. We now create different data sets for each temperature and test the effect of `species_genotype` within each temperature.
 
@@
   theme(legend.position = c(1, 5), axis.text.x = element_text(angle = 0, hjust = 0.5, size = 7), legend.text = element_text(size = 12, face = "italic"), axis.title = element_text(size = 12)) +
-  labs(x = "", y = "N. of macroconidia (x10.000)", color = "Species") +
+  labs(x = "", y = expression(N.~of~macroconia~x~10 ^ {
+    3
+  }), color = "Species") +
   ylim(0, 5) +
   guides(colour = guide_legend(nrow = 2))
+spor2
 ```
@@
 lmer_spor <- lmer(spores_ml ~ genotype + (1 | genotype / isolate), data = spor, REML = FALSE)
-
 ```
@@
 grid5 <- plot_grid(mgr1, grid1, labels = c("A"), rel_heights = c(1, 1), ncol = 1, align = "hv")
-ggsave(here("figs", "figure2.png"), grid5, width = 6, height = 6.2)
+ggsave(here("figs", "figure2.png"), grid5, dpi = 600, width = 6, height = 6.2)
 
@@
 The pathogenicity was assessed based on the progress of the symptoms from the central (inoculated) spikelet.
-The data was transformed to percent severity of the spike (proportion of diseased spikelets).
-There is one obsevation of severity for each day after inoculation (dai) and individual spike.
+The data were transformed to percent severity of the spike (from the proportion of diseased spikelets).
+There is one observation of severity for each day after inoculation (dai) on individual spike.
 
@@
   filter(isolate != "test") %>%
-  select (-c(8:12))
+  select(-c(8:12))
 
@@
 
-### Visualize temporal
+### Visualize temporal progress
 
@@
   labs(
+    shape = "Genotype",
+    color = "Species",
     y = "FHB severity (%) ", x = "Days after inoculation",
@@
     aes(min = mean_sev - sd_sev, max = mean_sev + sd_sev),
-    width = 0.2, alpha = 0.3, position = position_dodge(width = 0.9)
+    width = 0.2,
+    alpha = 0.3,
+    position = position_dodge(width = 0.9)
   ) +
@@
   labs(
-    y = "FHB severity (%) ", x = "Days after inoculation",
+    shape = "Genotype",
+    color = "Species",
+    y = "FHB severity (%) ",
+    x = "Days after inoculation",
     color = "species"
@@
 fig3 <- plot_grid(brsgua, brs194, labels = c("A", "B"), rel_widths = c(0.7, 1), ncol = 2, align = "hv")
-ggsave(here("figs", "figure3.png"), fig3, width = 9, height = 4)
+ggsave(here("figs", "figure3.png"), fig3, dpi = 600, width = 9, height = 4)
 
@@
 First, we need to group several variables up to `dai` (days after inoculation).
-To make our life easier, lets use the `do` in conjuction with the `tidy` functions of the `dplyr` and `broom` package, respectively.
+To make our life easier, lets use the `do` in conjunction with the `tidy` functions of the `dplyr` and `broom` package, respectively.
 This will calculate the audpc for each spike.
@@
 We fit a separate model for each cultivar because the experiments were conducted at different times.
-We fit a multilevel to test the effect of specis_genotype as fixed effects and spikes nested within isolates as random effects.
+We fit a multilevel to test the effect of `species_genotype` as fixed effects and spikes nested within isolates as random effects.
 The species_genotype interaction was non-significant (not shown), and then a simpler model is used.
@@
 mix_gua <- lmer(
-  audpc ~ species_genotype  +
-    (1 | isolate/spike),
-  data = audpc_gua, REML = FALSE)
+  audpc ~ genotype +
+    (1 | isolate / spike),
+  data = audpc_gua, REML = FALSE
+)
 
 mix_gua
-
 ```
@@
 Anova(mix_gua)
-means <- emmeans(mix_gua, ~ species_genotype)
+means <- emmeans(mix_gua, ~ genotype)
 cld(means)
@@
 ```{r}
-mix_194 <- lmer(audpc ~ species_genotype + 
-                  (1 | isolate/spike), 
-                data = audpc_194)
+mix_194 <- lmer(
+  audpc ~ genotype +
+    (1 | isolate / spike),
+  data = audpc_194
+)
 mix_194
@@
 Anova(mix_194)
-means <- emmeans(mix_194, ~ species_genotype)
+means <- emmeans(mix_194, ~ genotype)
 cld(means)
@@
   geom_point() +
-  geom_errorbar(aes(min = mean_ec50 - sd_ec50, max = mean_ec50 + sd_ec50), width = 0.2)+
+  geom_errorbar(aes(min = mean_ec50 - sd_ec50, max = mean_ec50 + sd_ec50), width = 0.2) +
   theme_few() +
-  coord_flip()+
-  theme(legend.position = "right", legend.text = element_text(size = 7, face = "italic")) +
+  coord_flip() +
+  theme(
+    legend.position = "right", axis.text.y = element_blank(),
+    legend.text = element_text(size = 7, face = "italic")
+  ) +
   ylim(0, 1.8) +
   scale_color_viridis(discrete = TRUE) +
-  labs(y = (expression(paste("EC"[50], " (", mu, "g/ml)"))), x = "FGSC strain", shape = "Species", color = "Genotype" ) +
-  ggsave(here("figs", "ec50-new.png"), width = 4.5, height = 3)
+  labs(y = (expression(paste("EC"[50], " (", mu, "g/ml)"))), x = "FGSC strain", shape = "Species", color = "Genotype") +
+  ggsave(here("figs", "figure4.png"), dpi = 600, width = 4.5, height = 3)
 ```

### 2018-01-21 data.Rmd
Check spell and text edit

@@
 The data for this study were obtained from five different (mostly repeated in time) experiments, and were organized in a single `xlsx` file.
-Data from each experiment are in a separate sheet and were exported to a `csv` file.
+Data from each experiment are in a separate worksheet and were exported to a `csv` file.
 They are already organized in a rectangular (tidy) format, mostly in the long-format (a raw is an observation).
-For some experiments, they are in the wide format (observations in multiple colums), and so will need some wrangling.
-This munging is done within the respective section of the data analysis page, after data import.
+For some experiments, they are in the wide format (observations in multiple columns), so some wrangling will be needed, which is shown within the respective section of the data analysis page, after data import.
 
@@
 
-> The data are stored in the following repositories:
-
-https://github.com/phytopathology/paper-FGSC-fitness/tree/master/dat  
+> The data are stored permanently at the Open Science Framework servers
 https://osf.io/c2mbr/ 
 Identifier: DOI 10.17605/OSF.IO/C2MBR
+
+>Citation:
+Del Ponte, E. M. (2018, January 21). Fitness of **Fusarium graminearum** species complex strains from Brazilian wheat. http://doi.org/10.17605/OSF.IO/C2MBR
